---
title: "flextable R package"
output:
  github_document:
    toc: true 
    toc_depth: 2
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Build Status](https://travis-ci.org/davidgohel/flextable.svg?branch=master)](https://travis-ci.org/davidgohel/flextable)
[![Coverage Status](https://img.shields.io/codecov/c/github/davidgohel/flextable/master.svg)](https://codecov.io/github/davidgohel/flextable?branch=master)
[![CRAN version](http://www.r-pkg.org/badges/version/flextable)](https://cran.r-project.org/package=flextable) ![](http://cranlogs.r-pkg.org/badges/grand-total/flextable)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


## `flextable` API


A tabular reporting tool that let R users create tables for reporting and control their formatting properties and their layout. 

Features:

  - Manage headers
  - Format text, paragraphs, cells and border of any element
  - Separate data from displayed values
  - Outut availables are Microsoft Word, PowerPoint and HTML (as an `htmlwidgets` object).

## Install `flextable`

You can get the development version from GitHub:

```{r eval = FALSE}
devtools::install_github("davidgohel/flextable")
```


## Load `flextable`

`flextable` is designed for use with the `%>%` pipe operator. You can make this `%>%` operator available in your R workspace by loading package [`magrittr`](https://cran.r-project.org/package=magrittr).

```{r load-package}
library("flextable")
library("magrittr")
```


## Quick demo

First, load packages.

```{r warning=FALSE, message=FALSE}
library(flextable)
library(oxbase)
library(dplyr)
```

We will use the following dataset:

```{r}
mydata <- iris %>% 
  group_by(Species) %>% 
  do( head(., n = 3) )
```


Then defined few styles to be used later.

```{r}
def_txt <- fp_text(font.size = 9, font.family = "Arial", color = "gray40")
def_head <- update(def_txt, font.size = 11, color = "gray20")
def_cell <- fp_cell( border = fp_border(color = "gray") )
```

Define the flextable and its layout (headers, merge cells, etc.).

```{r}
ft <- flextable(mydata, col_keys = c("Species", 
    "Sepal.Length", "Sepal.Width", "Petal.Length", 
    "Petal.Width") ) %>%
  set_header_labels(Sepal.Length = "Length", 
             Sepal.Width = "Width",
             Petal.Length = "Length",
             Petal.Width = "Width", 
             Species = "Species" ) %>%
  merge_v(j = "Species" ) 
```

Then apply some formatting properties :

```{r 'tabwid_example'}
ft %>% 
  theme_tron() %>% 
  autofit() %>% 
  tabwid()
```


