---
title: "flextable R package"
output:
  github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Build Status](https://travis-ci.org/davidgohel/flextable.svg?branch=master)](https://travis-ci.org/davidgohel/flextable)
[![Coverage Status](https://img.shields.io/codecov/c/github/davidgohel/flextable/master.svg)](https://codecov.io/github/davidgohel/flextable?branch=master)
[![CRAN version](http://www.r-pkg.org/badges/version/flextable)](https://cran.r-project.org/package=flextable) ![](http://cranlogs.r-pkg.org/badges/grand-total/flextable)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```


## flextable

The flextable package provides a framework for easily creating tables for reporting. 
Tables created using the framework can be:

* Embedded within R Markdown documents and Shiny web applications.
* Embedded within Microsoft Word or PowerPoint documents.

## Getting Started

An API is available to let R users create tables for reporting and control their formatting properties and their layout. A `flextable` object is a data.frame representation, it can be manipulated with functions that give control over:

- headers managment 
- text, paragraphs, cells and border formatting of any element
- displayed values

### Installation

You can get the development version from GitHub:

```{r eval = FALSE}
devtools::install_github("davidgohel/flextable")
```

There are articles on the flextable website that will help you get you to start quickly:

* [Introduction to flextable](http://davidgohel.github.io/flextable/articles/introduction.html)
* [Function reference](http://davidgohel.github.io/flextable/reference/index.html)




