---
title: "flextable R package"
output:
  github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Build Status](https://travis-ci.org/davidgohel/flextable.svg?branch=master)](https://travis-ci.org/davidgohel/flextable)
[![Build status](https://ci.appveyor.com/api/projects/status/github/davidgohel/flextable?branch=master)](https://ci.appveyor.com/project/davidgohel/flextable/branch/master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/flextable)](https://cran.r-project.org/package=flextable)
![cranlogs](http://cranlogs.r-pkg.org./badges/flextable)
[![codecov](https://codecov.io/gh/davidgohel/flextable/branch/master/graph/badge.svg)](https://codecov.io/gh/davidgohel/flextable)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

<img src="http://www.ardata.fr/images/flextablelogo.svg" style="width:15%;"/>

```{r, echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```



The flextable package provides a framework for easily create tables for reporting. 
Tables can be embedded within:

* R Markdown documents
* Microsoft Word or PowerPoint documents.


## Getting Started

An API is available to let R users create tables for reporting and control their formatting properties and their layout. A `flextable` object is a data.frame representation, it can be manipulated with functions that give control over:

- headers content 
- text, paragraphs, cells and border formatting of any element
- displayed values

There are articles on the flextable website that will help you get you to start quickly:

* [Introduction to flextable](https://davidgohel.github.io/flextable/articles/overview.html)
* [Function reference](https://davidgohel.github.io/flextable/reference/index.html)

The help pages can be read [here](https://davidgohel.github.io/flextable).

### Installation


```{r eval = FALSE}
install.packages("flextable")
```

You can get the development version from GitHub:

```{r eval = FALSE}
devtools::install_github("davidgohel/flextable")
```


## Ressources

### Online documentation

The help pages are located at https://davidgohel.github.io/flextable/.

### Getting help

This project is developped and maintained on my own time. If you 
have questions about how to use the package, visit
Stackoverflow and use tags `flextable` and `r` [Stackoverflow
link](https://stackoverflow.com/questions/tagged/flextable+r)\! I
usually read them and answer when possible.

## Contributing to the package

### Code of Conduct

Anyone getting involved in this package agrees to our [Code of Conduct](https://github.com/davidgohel/flextable/blob/master/CONDUCT.md).

### Bug reports

When you file a [bug report](https://github.com/davidgohel/flextable/issues), please spend some time making it easy for me to follow and reproduce. The more time you spend on making the bug report coherent, the more time I can dedicate to investigate the bug as opposed to the bug report.

### Contributing to the package development

A great way to start is to contribute an example or improve the documentation.

If you want to submit a Pull Request to integrate functions of yours, please provide:

* the new function(s) with code and roxygen tags (with examples)
* a new section in the appropriate vignette that describes how to use the new function
* corresponding tests in directory `tests/testthat`.

By using rhub (run `rhub::check_for_cran()`), you will see if everything is ok.
When submitted, the PR will be evaluated automatically on travis and appveyor and you will be able to see if something broke.

