---
title: "flextable R package"
output:
  github_document:
    toc: true 
    toc_depth: 3
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Build Status](https://travis-ci.org/davidgohel/flextable.svg?branch=master)](https://travis-ci.org/davidgohel/flextable) [![Coverage Status](https://coveralls.io/repos/davidgohel/flextable/badge.svg)](https://coveralls.io/r/davidgohel/flextable) [![CRAN version](http://www.r-pkg.org/badges/version/flextable)](https://cran.r-project.org/package=flextable) ![](http://cranlogs.r-pkg.org/badges/grand-total/flextable)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


`flextable` API
---

A tabular reporting tool that let R users create tables for reporting and control their formatting properties and their layout. 

Features:

  - Manage headers
  - Format text, paragraphs, cells and border of any element
  - Separate data from displayed values
  - Outut availables are Microsoft Word, PowerPoint and HTML (as an `htmlwidgets` object).


### Install `flextable`

You can get the development version from GitHub:

```{r eval = FALSE}
devtools::install_github("davidgohel/flextable")
```


### Load `flextable`

`flextable` is designed for use with the `%>%` pipe operator. You can make this `%>%` operator available in your R workspace by loading package [`magrittr`](https://cran.r-project.org/package=magrittr).

```{r load-package}
library("flextable")
library("magrittr")
```


### Quick demo

TBD
