---
title: "flextable display value"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Flextable display value}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```


```{r warning=FALSE, echo=FALSE, message=FALSE}
library(officer)
library(flextable)
```


By default, the displayed content of each cell will be the 
result of a simple formatting. The content can also be 
*comoosed* as the result of a concatenation of several *chunks*.

There are `formatter` functions to define how columns content will be 
display (e.g. to define the number of decimal). This is 
a column-wise operation.

User can also have more control over displayed content by using function `compose`.
The function enable to define the elements composing the paragraph and their respective formats.
It can also be used to mix text chunks and images.

## Column-wise functions

Theses are high level functions that should satisfy most 
of the usual needs. They can be used to define the formatting 
of one or more columns.

Each accept a `prefix` and `suffix` argument that can be used 
to add a currency symbol for example. Also they all have 
`na_str` argument (defaut to ""), the string to use when data 
are not available.

- `colformat_num()` with arguments `digits` and `big.mark`
- `colformat_int()` with arguments `big.mark`
- `colformat_char()`

```{r}
dat <- head(mtcars, n = 10)
dat[3:7, 1] <- NA
dat[, 2] <- dat[, 6] * 1000000

ft <- flextable(dat)
num_keys <- c("mpg", "disp", "drat", "wt", "qsec")
int_keys <- c("cyl", "hp", "vs", "am", "gear", "carb")

ft <- colformat_num(x = ft, col_keys = num_keys, big.mark = ",", digits = 2, na_str = "missing")
ft <- colformat_int(x = ft, col_keys = int_keys, big.mark = ",")
autofit(ft)
```


## set_formatter function

It is possible to define the functions that will be used to format the data.frame 
values into strings. `set_formatter` set column formatter functions. 


```{r}
ft <- flextable(head(mtcars, n = 10 ), 
                   col_keys = c("gear", "mpg", "qsec"))
ft <- set_formatter(ft, 
    mpg = function(x) sprintf("%.04f", x),
    gear = function(x) sprintf("%.0f gears", x)
  )
ft <- theme_booktabs(ft)
ft <- autofit(ft)
ft
```


## compose function

`flextable` content can be defined with function `compose`.

It lets user control the formated content at the cell level of the table. 
It is possible to define a content for a row subset and a column as well 
as on the whole column. One can mix images and text (but not with 
PowerPoint because PowerPoint can not do it).

Below the starting point of next illustrations:

```{r}
myft <- flextable( head(mtcars), 
  col_keys = c("am", "separator", "gear", "mpg", "drat", "qsec" ))
myft <- bold(myft, part = "header")
myft <- border(myft, border = fp_border( width = 0), 
  border.top = fp_border(), border.bottom = fp_border(), 
  part = "all")
myft <- align(myft, align = "right", part = "all" )
myft <- border(myft, j = ~ separator, border = fp_border(width=0), part = "all")
myft <- width(myft, j = ~ separator, width = .1)
myft
```

The function requires a call to `as_paragraph` which will concatenate text or images chunks as 
a paragraph. 

![](https://www.ardata.fr/img/illustrations/flextable_compose.svg){width=350px}


The following example shows how to control the format of displayed values and how to 
associate them with specific text formatting properties (bold red text):

```{r}
myft <- compose( 
  myft, j = "mpg", 
  value = as_paragraph(
    "mpg value is ", 
    as_chunk(sprintf("%.01f", mpg), props = fp_text(color = "red", bold = TRUE) ) )
  )
myft <- autofit(myft)
myft
```

With that system, it's easy to concatenate multiple values:

```{r}
myft <- compose( 
  myft, j = "mpg", 
  value = as_paragraph(
    "mpg value is ", 
    as_chunk(sprintf("%.01f", mpg), props = fp_text(color = "red", bold = TRUE) ), 
    " with ",
    as_chunk(sprintf("# %.0f", carb), props = fp_text(color = "gray", italic = TRUE) )
    )
  )

myft <- autofit(myft)
myft
```

Or to define specific title headers:

```{r}
myft <- compose( 
  myft, j = "mpg", part = "header",
  value = as_paragraph(
    "Miles/(US) gallon ", 
    as_chunk("* with num of carb.", props = fp_text(color = "gray", vertical.align = "superscript") )
    )
  )

myft <- autofit(myft)
myft
```

### Images

Function `compose` supports images insertion. Use function `as_image` in `as_paragraph` call:

```{r}
img.file <- file.path( R.home("doc"), "html", "logo.jpg" )

myft <- compose( myft, i = ~ qsec > 18, j = "qsec", 
  value = as_paragraph(as_image( src = img.file, width = .20, height = .15))
)
myft <- autofit(myft)
myft
```

### minibar

Function `compose` supports *mini barplots* insertion. Use function `minibar` in `as_paragraph` call:

```{r}
myft <- flextable( head(iris, n = 10 ))

myft <- compose( myft, j = 1,
  value = as_paragraph(
    minibar(value = Sepal.Length, max = max(Sepal.Length))
  ),
  part = "body")

autofit(myft)
```


### linerange

Function `compose` supports *mini linerange* insertion. Use function `linerange` in `as_paragraph` call:

```{r}
myft <- flextable( head(iris, n = 10 ))

myft <- compose( myft, j = 1,
  value = as_paragraph(
    linerange(value = Sepal.Length, max = max(Sepal.Length))
  ),
  part = "body")

autofit(myft)
```



