---
title: "Complex text formatting"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Complex text formatting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r warning=FALSE, echo=FALSE}
library(flextable)
library(oxbase)
library(dplyr)
```

Below the starting point of next illustrations:

```{r}
myft <- flextable(head(mtcars), 
                  col_keys = c("am", "separator", "gear", "mpg", "drat" )) %>% 
  bold(part = "header") %>% 
  style(pr_c = c_b_b(), pr_p = fp_right(padding=2), part = "all" ) %>%
  border(j = ~ separator, border = fp_border(width=0), part = "all")

tabwid(myft)
```

## display function

Flextable content is defined with `display` function. 

It always requires a call to `fpar` which is a simple wrapper whose role is to concatenate  formated text. The following call to fpar only let you modify the display content of a cell (because fpar argument is only a single string). 

```{r}
myft <- myft %>%
  display( 
    am = fpar(formatC(am, format = "f", digits = 0 ) ) 
  )

tabwid(myft)
```

If more control is needed, i.e. for conditionnal formatting, one can use function `ftext` to associate a text with formatting properties.

```{r}
myft <- myft %>%
  display(i = ~ drat > 3.6, 
    gear = fpar( ftext( gear, prop = fp_bold(color="red") ) )
  )

tabwid(myft)
```

`fpar` can have multiple arguments that will be then concatenated.

```{r}
myft <- myft %>%
  display(
    mpg = fpar(mpg, " and ", ftext( carb, prop = fp_bold()), " carb(s)." ) 
  )

tabwid(myft)
```

`fpar` handle call to `minibar`. This will generate a simple bar to be inserted as content.

```{r}
myft <- myft %>%
  display( i = ~ gear < 4,
    drat = fpar(minibar( value = drat, max = max(.$drat), 
                         barcol = "#C90000", width = 1, height = .15) )
  ) %>% autofit()

tabwid(myft)
```

Note usage of `.`, this object represents the input dataset of `flextable` call. The expression is not vectorized 


