---
title: "Format flextable objects"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Format flextable objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>", 
  eval = !is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))
)

dir.create("assets/docx", recursive = TRUE, showWarnings = FALSE)
dir.create("assets/pptx", recursive = TRUE, showWarnings = FALSE)
office_doc_link <- function(url){
  stopifnot(requireNamespace("htmltools", quietly = TRUE))
  htmltools::tags$p(  htmltools::tags$span("Download file "),
    htmltools::tags$a(basename(url), href = url), 
    htmltools::tags$span(" - view with"),
    htmltools::tags$a("office web viewer", target="_blank", 
      href = paste0("https://view.officeapps.live.com/op/view.aspx?src=", url)
      ), 
    style="text-align:center;font-style:italic;color:gray;"
    )
}
```


```{r warning=FALSE, echo=FALSE, message=FALSE}
library(officer)
library(flextable)
```

A flextable is made of two parts, an header and a body. **To specify which part formatting instructions should affect, use argument `part`.** Possible values are:

* "header": the header part of the table
* "body": the body part of the table
* "all": the body and the header parts of the table


## Shortcuts functions

There are simple functions to modify formatting properties of flextable objects: `bg`, `bold`, `border`, `color`, `padding`, `fontsize`, `italic`, `align`.

```{r}
myft <- regulartable(head(iris))
myft
```

### bold

```{r}
myft <- regulartable(head(iris)) 
myft <- bold(myft, part = "header") # bold header
myft
```

### Font size

```{r}
myft <- fontsize(myft, part = "header", size = 12) 
myft
```

### change font color

```{r}
myft <- color(myft, color = "#E4C994")
myft
```

### Italic

```{r}
myft <- italic(myft, i = ~ Sepal.Length > 5, 
         j = ~ Sepal.Length + Sepal.Width, italic = TRUE)
myft
```

### change background color

```{r}
# light gray as background color for header
myft <-  bg(myft, bg = "#E4C994", part = "header")
# dark gray as background color for body
myft <-  bg(myft, bg = "#333333", part = "body")
myft
```

### Text alignment

```{r}
myft <- align( myft, align = "center", part = "all" )
myft
```


### add padding

```{r}
myft <- padding( myft, padding = 3, part = "all" )
myft
```

### change border

Borders properties can be defined with functions:

* `hline()`: set bottom borders of one or all parts of a flextable. 
* `hline_top()`: set top border of one or all parts of a flextable. 
* `hline_bottom()`: set bottom border of one or all parts of a flextable. 
* `vline()`: set right borders of one or all parts of a flextable. 
* `vline_left()`: set left border of one or all parts of a flextable.
* `vline_right()`: set right border of one or all parts of a flextable.

```{r}
# remove all defined borders
myft <- border( myft, border = fp_border(width=0), part = "all" )

big_b <- fp_border(color="gray70", width = 2)
std_b <- fp_border(color="white")

myft <- vline( myft, border = std_b, part = "all" )
myft <- vline_left( myft, border = big_b, part = "all" )
myft <- vline_right( myft, border = big_b, part = "all" )
myft <- hline( myft, border = std_b )
myft <- hline_bottom( myft, border = big_b )
myft <- hline_top( myft, border = big_b, part = "all" )
myft
```

### Text rotation 

Text rotation is possible in flextable objects but will only work with Word and PowerPoint outputs. This is achieved by using function `rotate`.

Argument `rotation` is mandatory and expects one of these values:

- "lrtb", left to right top to bottom
- "tbrl", top to bottom right to left
- "btlr", bottom to top left to right

Argument `align` is used for cell content vertical alignment, it should be one of these values: "top", "bottom" or "center".

```{r}
ft <- regulartable(head(iris))
ft <- rotate(ft, rotation = "tbrl", align = "top", part = "header")
ft <- theme_vanilla(ft)
ft <- autofit(ft)

# as autofit do not handle rotation, you will have
# to change manually header cells'height.
ft <- height(ft, height = 1, part = "header")
ft
```

The following example is producing a Word document:

```{r results='hide'}
library(officer)
doc <- read_docx()
doc <- body_add_flextable(doc, ft)
print(doc, target = "assets/docx/rotate.docx")
```


```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/flextable/articles/", "assets/docx/rotate.docx" ) )
```

The following example is producing a PowerPoint document:

```{r results='hide'}
library(officer)
doc <- read_pptx()
doc <- add_slide(doc, layout = "Title and Content", master = "Office Theme")
doc <- ph_with_flextable(doc, ft)
print(doc, target = "assets/pptx/rotate.pptx")
```


```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/flextable/articles/", "assets/pptx/rotate.pptx" ) )
```

## Conditional formatting

Conditional formatting can be made by using the selector arguments.

```{r warning=FALSE, message=FALSE}
myft <- color(myft, i = ~ Sepal.Length < 5 & Petal.Length > 1.3, 
        j = ~ Petal.Width + Species, 
        color="red")
myft <- bg(myft, j = 1, bg = "#D3C994", part = "header")
myft <- italic(myft, i = ~ Sepal.Length > 5)
myft <- bold(myft, i = 4, j = "Sepal.Length")
myft
```

`i` and `j` arguments can be also standard R vectors:

```{r warning=FALSE, message=FALSE}
row_id <- with(head(iris), Sepal.Length < 5 & Petal.Length > 1.3 )
col_id <- c("Petal.Width", "Species")

myft <- color(myft, i = row_id, j = col_id, color="red") 

myft
```

## Function style

The `style` function lets you style a selection of the flextable with several formatting properties. 

> It main advantage is to let specify a set of formatting properties for a selection. 

Package officer needs to be loaded, it provides the following formatting properties:

* for text: `fp_text`
* for paragraphs: `fp_par`
* for cells: `fp_cell` and `fp_border`


```{r}
library(officer)
def_cell <- fp_cell(border = fp_border(color="#00C9C9"))
def_par <- fp_par(text.align = "center")
def_text <- fp_text(color="#999999", italic = TRUE)
def_text_header <- update(color="black", def_text, bold = TRUE)

ft <- regulartable(head(mtcars, n = 10 ))
ft <- style( ft, pr_c = def_cell, pr_p = def_par, pr_t = def_text, part = "all")  
ft

ft <- style( ft, pr_t = def_text_header, part = "header")  
ft
```


## set_formatter function

When working with `regulartable`, it is possible to define the functions that will be used to format the data.frame values into strings. `set_formatter` set column formatter functions.

> Note `set_formatter` only works with `regulartable` objects, use `display` for `flextable` objects.

```{r}
ft <- regulartable(head(mtcars, n = 10 ), 
                   col_keys = c("gear", "mpg", "qsec"))
ft <- set_formatter(ft, 
    mpg = function(x) sprintf("%.04f", x),
    gear = function(x) sprintf("%.0f gears", x)
  )
ft <- theme_booktabs(ft)
ft <- autofit(ft)
ft
```


## display function

`flextable` content is defined with `display` function. 

> Note `display` only works with `flextable` objects, use `set_formatter` for `regulartable` objects.

Below the starting point of next illustrations:

```{r}
myft <- flextable( head(mtcars), 
  col_keys = c("am", "separator", "gear", "mpg", "drat", "qsec" ))
myft <- bold(myft, part = "header")
myft <- border(myft, border = fp_border( width = 0), 
  border.top = fp_border(), border.bottom = fp_border(), 
  part = "all")
myft <- align(myft, align = "right", part = "all" )
myft <- border(myft, j = ~ separator, border = fp_border(width=0), part = "all")
myft <- width(myft, j = ~ separator, width = .1)
myft
```


The function requires argument `pattern` which is a string template inspired by *[mustaches](http://mustache.github.com/)*. The string will be expanded with tags using values provided in `formatters` argument; tags can eventually be formatted with `fprops` argument. 

The following example shows how to control the format of displayed values and how to associate them with specific text formatting properties (bold red text):

```{r}
myft <- display( myft, col_key = "mpg", pattern = "{{mpg}}", 
    formatters = list(mpg ~ sprintf("%.01f", mpg) ), 
    fprops = list(mpg = fp_text(color = "red", italic = TRUE) )
  )

myft
```

With that system, it's easy to concatenate multiple values:

```{r}
myft <- display( myft, i = ~ drat > 3.6, 
           col_key = "mpg", pattern = "{{mpg}} with {{carb}}", 
           formatters = list(mpg ~ sprintf("%.01f", mpg), 
                             carb ~ sprintf("# %.0f carb.", carb) ), 
              fprops = list(mpg = fp_text(color = "#CC55CC", bold = TRUE) )
  )
myft <- autofit(myft)
myft
```

Or to define specific title headers:

```{r}
myft <- display( myft, col_key = "mpg", 
   part = "header",
   pattern = "Miles/(US) gallon {{my_message}}", 
   formatters = list(
     my_message ~ sprintf("* with num of carb.") 
     ), 
   fprops = list(
     my_message = fp_text(color = "gray", vertical.align = "superscript")
     ) 
   )
myft <- autofit(myft)
myft
```

### Images

Function `display` supports images insertion. Use function `as_image` within `formatters` argument.

```{r}
img.file <- file.path( R.home("doc"), "html", "logo.jpg" )

myft <- display( myft, i = ~ qsec > 18, col_key = "qsec", 
           pattern = "blah blah {{r_logo}} {{qsec}}",
           formatters = list(
             r_logo ~ as_image(qsec, src = img.file, width = .20, height = .15), 
             qsec ~ sprintf("qsec: %.1f", qsec) ), 
           fprops = list(qsec = fp_text(color = "orange", vertical.align = "superscript"))
           )
myft <- autofit(myft)
myft
```
