---
title: "Format flextable objects"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Format flextable objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)

dir.create("assets/format", recursive = TRUE, showWarnings = FALSE)
office_doc_link <- function(url){
  stopifnot(requireNamespace("htmltools", quietly = TRUE))
  htmltools::tags$p(  htmltools::tags$span("Download file "),
    htmltools::tags$a(basename(url), href = url), 
    htmltools::tags$span(" - view with"),
    htmltools::tags$a("office web viewer", target="_blank", 
      href = paste0("https://view.officeapps.live.com/op/view.aspx?src=", url)
      ), 
    style="text-align:center;font-style:italic;color:gray;"
    )
}
```


```{r warning=FALSE, echo=FALSE, message=FALSE}
library(flextable)
library(dplyr)
```

A flextable is made of two parts, an header and a body. **To specify which part formatting instructions should affect, use argument `part`.** Possible values are:

* "header": the header part of the table
* "body": the body part of the table
* "all": the body and the header parts of the table


## Shortcuts functions

There are simple functions to modify formatting properties of flextable objects: `bg`, `bold`, `border`, `color`, `padding`, `fontsize`, `italic`, `align`.

```{r}
myft <- flextable(head(iris))
tabwid(myft)
```

### bold

```{r}
myft <- flextable(head(iris)) %>% 
  # bold header
  bold(part = "header") 
tabwid(myft)
```

### Font size

```{r}
myft <- myft %>% fontsize(part = "header", size = 12) 
tabwid(myft)
```

### change font color

```{r}
myft <- myft %>% color(color = "#E4C994")
tabwid(myft)
```

### Italic

```{r}
myft <- myft %>% 
  italic(i = ~ Sepal.Length > 5, 
         j = ~ Sepal.Length + Sepal.Width, italic = TRUE)
tabwid(myft)
```

### change background color

```{r}
myft <- myft %>% 
  # light gray as background color for header
  bg(bg = "#E4C994", part = "header") %>% 
  # dark gray as background color for body
  bg(bg = "#333333", part = "body")

tabwid(myft)
```

### Text alignment

```{r}
myft <- myft %>% align( align = "center", part = "all" )
tabwid(myft)
```


### add padding

```{r}
myft <- myft %>% padding( padding = 3, part = "all" )
tabwid(myft)
```

### change border

```{r}
myft <- myft %>% 
  border( border = fp_border(color="white"), part = "all" )
  
tabwid(myft)
```


## Conditional formatting

Conditional formatting can be made by using the selector arguments.

```{r warning=FALSE, message=FALSE}
myft <- myft %>% 
  color(i = ~ Sepal.Length < 5 & Petal.Length > 1.3, 
        j = ~ Petal.Width + Species, 
        color="red") 
tabwid(myft)
```

`i` and `j` arguments can be also standard R vectors:

```{r warning=FALSE, message=FALSE}
row_id <- with(head(iris), Sepal.Length < 5 & Petal.Length > 1.3 )
col_id <- c("Petal.Width", "Species")

myft <- color(myft, i = row_id, j = col_id, color="red") 

tabwid(myft)
```

## Function style

The `style` function lets you style a selection of the flextable with several formatting properties. 

> It main advantage is to let specify a set of formatting properties for a selection. 

Package officer needs to be loaded, it provides the following formatting properties:

* for text: `fp_text`
* for paragraphs: `fp_par`
* for cells: `fp_cell` and `fp_border`


```{r}
library(officer)
def_cell <- fp_cell(border = fp_border(color="#00C9C9"))
def_par <- fp_par(text.align = "center")
def_text <- fp_text(color="#999999", italic = TRUE)
def_text_header <- update(color="black", def_text, bold = TRUE)

ft <- flextable(head(mtcars, n = 10 )) %>% 
  style( pr_c = def_cell, pr_p = def_par, pr_t = def_text, part = "body")  
tabwid(ft)

ft <- ft %>% 
  style( pr_c = def_cell, pr_p = def_par, pr_t = def_text_header, part = "header")  
tabwid(ft)
```

### Text rotation 

Text rotation is possible in flextable objects but will only work with Word and PowerPoint outputs. This is achieved by using function `rotate`.

Argument `rotation` is mandatory and expects one of these values:

- "lrtb", left to right top to bottom
- "tbrl", top to bottom right to left
- "btlr", bottom to top left to right

Argument `align` is used for cell content vertical alignment, it should be one of these values: "top", "bottom" or "center".

```{r}
ft <- flextable(head(iris)) %>% 
  rotate(rotation = "tbrl", align = "top", part = "header") %>% 
  theme_vanilla() %>% 
  autofit() %>% 
  # as autofit do not handle rotation, you will have
  # to change manually header cells'height.
  height(height = 1, part = "header")
```

```{r}
library(officer)
read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "assets/format/rotate.docx") %>% 
  invisible()
```


```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/flextable/articles/", "assets/format/rotate.docx" ) )
```

## display function

Below the starting point of next illustrations:

```{r}
myft <- flextable( head(mtcars), 
  col_keys = c("am", "separator", "gear", "mpg", "drat", "qsec" )) %>% 
  bold(part = "header") %>% 
  border(border = fp_border( width = 0), 
         border.top = fp_border(), border.bottom = fp_border(), 
         part = "all") %>% 
  align(align = "right", part = "all" ) %>%
  border(j = ~ separator, border = fp_border(width=0), part = "all") %>% 
  width(j = ~ separator, width = .1)

tabwid(myft)
```

Flextable content is defined with `display` function. The function requires argument `pattern` which is a string template inspired by *[mustaches](http://mustache.github.com/)*. The string will be expanded with tags using values provided in `formatters` argument; tags can eventually be formatted with `fprops` argument. 

The following example shows how to control the format of displayed values and how to associate them with specific text formatting properties (bold red text):

```{r}
myft <- myft %>%
  display( col_key = "mpg", pattern = "{{mpg}}", 
           formatters = list(mpg ~ sprintf("%.01f", mpg) ), 
              fprops = list(mpg = fp_text(color = "red", italic = TRUE) )
  )

tabwid(myft)
```

With that system, it's easy to concatenate multiple values:

```{r}
myft <- myft %>%
  display( i = ~ drat > 3.6, 
           col_key = "mpg", pattern = "{{mpg}} with {{carb}}", 
           formatters = list(mpg ~ sprintf("%.01f", mpg), 
                             carb ~ sprintf("# %.0f carb.", carb) ), 
              fprops = list(mpg = fp_text(color = "#CC55CC", bold = TRUE) )
  ) %>% autofit()

tabwid(myft)
```

Or to define specific title headers:

```{r}
myft <- myft %>%
  display( col_key = "mpg", pattern = "{{mpg}} {{my_message}}", part = "header",
           formatters = list(mpg ~ "Miles/(US) gallon", 
                             my_message ~ sprintf("* with num of carb.") ), 
              fprops = list(my_message = fp_text(color = "gray", vertical.align = "superscript")
                            )
  ) %>% autofit()

tabwid(myft)
```

### Images

Function `display` supports images insertion. Use function `as_image` within `formatters` argument.

```{r}
img.file <- file.path( Sys.getenv("R_HOME"), "doc", "html", "logo.jpg" )

myft <- myft %>%
  display( i = ~ qsec > 18, col_key = "qsec", 
           pattern = "blah blah {{r_logo}} {{qsec}}",
           formatters = list(
             r_logo ~ as_image(qsec, src = img.file, width = .20, height = .15), 
             qsec ~ sprintf("qsec: %.1f", qsec) ), 
           fprops = list(qsec = fp_text(color = "orange", vertical.align = "superscript"))
           ) %>% 
  autofit()

tabwid(myft)
```
