---
title: "Format flextable objects"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Format flextable objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>", 
  eval = !is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))
)

dir.create("assets/docx", recursive = TRUE, showWarnings = FALSE)
dir.create("assets/pptx", recursive = TRUE, showWarnings = FALSE)
office_doc_link <- function(url){
  stopifnot(requireNamespace("htmltools", quietly = TRUE))
  htmltools::tags$p(  htmltools::tags$span("Download file "),
    htmltools::tags$a(basename(url), href = url), 
    htmltools::tags$span(" - view with"),
    htmltools::tags$a("office web viewer", target="_blank", 
      href = paste0("https://view.officeapps.live.com/op/view.aspx?src=", url)
      ), 
    style="text-align:center;font-style:italic;color:gray;"
    )
}
```


```{r warning=FALSE, echo=FALSE, message=FALSE}
library(officer)
library(flextable)
```

A flextable is made of parts, an header, a body and a footer. **To specify 
which part formatting instructions should affect, use argument `part`.**
Possible values are:

* "header": the header part of the table
* "footer": the footer part of the table
* "body": the body part of the table
* "all": the body and the header parts of the table


## Shortcuts functions

There are simple functions to modify formatting properties of flextable objects: `bg`, `bold`, `border`, `color`, `padding`, `fontsize`, `italic`, `align`, ...

They are illustrated in the examples below.

```{r}
myft <- flextable(head(iris))
myft
```

### bold

```{r}
myft <- flextable(head(iris)) 
myft <- bold(myft, part = "header") # bold header
myft
```

### Font size

```{r}
myft <- fontsize(myft, part = "header", size = 12) 
myft
```

### change font color

```{r}
myft <- color(myft, color = "#E4C994")
myft
```

### Italic

```{r}
myft <- italic(myft, i = ~ Sepal.Length > 5, 
         j = ~ Sepal.Length + Sepal.Width, italic = TRUE)
myft
```

### change background color

```{r}
# light gray as background color for header
myft <-  bg(myft, bg = "#E4C994", part = "header")
# dark gray as background color for body
myft <-  bg(myft, bg = "#333333", part = "body")
myft
```

### Text alignment

```{r}
myft <- align( myft, align = "center", part = "all" )
myft
```


### add padding

```{r}
myft <- padding( myft, padding = 3, part = "all" )
myft
```

### change font

```{r}
myft <- font(myft, j = "Species", fontname = "Times")
myft <- fontsize(myft, j = "Species", size = 14)
myft
```

### Text rotation 

Text rotation is possible in flextable objects but will only work 
correctly with Word and PowerPoint outputs (html output is not 
always correct). This is achieved by using function `rotate()`.

Argument `rotation` is mandatory and expects one of these values:

- "lrtb", left to right top to bottom
- "tbrl", top to bottom right to left
- "btlr", bottom to top left to right

Argument `align` is used for cell content vertical alignment, it should be one of these values: "top", "bottom" or "center".

```{r}
ft <- flextable(head(iris))
ft <- rotate(ft, rotation = "tbrl", align = "top", part = "header")
ft <- theme_vanilla(ft)
ft <- autofit(ft)

# as autofit do not handle rotation, you will have
# to change manually header cells'height.
ft <- height(ft, height = 1, part = "header")
ft
```


## Borders

Borders properties can be defined with functions:

* `hline()`: set bottom borders of one or all parts of a flextable. 
* `hline_top()`: set top border of one or all parts of a flextable. 
* `hline_bottom()`: set bottom border of one or all parts of a flextable. 
* `vline()`: set right borders of one or all parts of a flextable. 
* `vline_left()`: set left border of one or all parts of a flextable.
* `vline_right()`: set right border of one or all parts of a flextable.

```{r}
# remove all defined borders
myft <- border_remove( myft )

big_b <- fp_border(color="gray70", width = 2)
std_b <- fp_border(color="white")

myft <- vline( myft, border = std_b, part = "all" )
myft <- vline_left( myft, border = big_b, part = "all" )
myft <- vline_right( myft, border = big_b, part = "all" )
myft <- hline( myft, border = std_b )
myft <- hline_bottom( myft, border = big_b )
myft <- hline_top( myft, border = big_b, part = "all" )
myft
```

Or when borders design is simple with:

* `border_outer()`: apply a border to outer cells of one or all parts 
of a flextable.
* `border_inner()`: apply a border to inner cells of one or all parts 
of a flextable.


```{r}
std_b2 <- fp_border(color="white", style = "dashed")

# remove all defined borders
myft <- border_remove( myft )

myft <- border_outer( myft, border = big_b, part = "all" )
myft <- border_inner_h( myft, border = std_b, part = "all" )
myft <- border_inner_v( myft, border = std_b2, part = "all" )
myft
```


## Conditional formatting

Conditional formatting can be made by using the selector arguments.

```{r warning=FALSE, message=FALSE}
myft <- color(myft, i = ~ Sepal.Length < 5 & Petal.Length > 1.3, 
        j = ~ Petal.Width + Species, 
        color="red")
myft <- bg(myft, j = 1, bg = "#D3C994", part = "header")
myft <- italic(myft, i = ~ Sepal.Length > 5)
myft <- bold(myft, i = 4, j = "Sepal.Length")
myft
```

`i` and `j` arguments can be also standard R vectors:

```{r warning=FALSE, message=FALSE}
row_id <- with(head(iris), Sepal.Length < 5 & Petal.Length > 1.3 )
col_id <- c("Petal.Width", "Species")

myft <- color(myft, i = row_id, j = col_id, color="red") 

myft
```

## Function style

The `style` function lets you style a selection of the flextable with several formatting properties. 

> It main advantage is to let specify a set of formatting properties for a selection. 

Package officer needs to be loaded, it provides the following formatting properties:

* for text: `fp_text`
* for paragraphs: `fp_par`
* for cells: `fp_cell` and `fp_border`


```{r}
library(officer)
def_cell <- fp_cell(border = fp_border(color="#00C9C9"))
def_par <- fp_par(text.align = "center")
def_text <- fp_text(color="#999999", italic = TRUE)
def_text_header <- update(color="black", def_text, bold = TRUE)

ft <- flextable(head(mtcars, n = 10 ))
ft <- style( ft, pr_c = def_cell, pr_p = def_par, pr_t = def_text, part = "all")  
ft

ft <- style( ft, pr_t = def_text_header, part = "header")  
ft
```


## Content format functions

By default, the displayed content of each cell will be the 
result of a simple formatting. The content can also be 
*comoosed* as the result of a concatenation of several *chunks*.

There are `formatter` functions to define how columns content will be 
display (e.g. to define the number of decimal). This is 
a column-wise operation.

User can also have more control over displayed content by using function `compose`.
The function enable to define the elements composing the paragraph and their respective formats.
It can also be used to mix text chunks and images.

### Column-wise functions

Theses are high level functions that should satisfy most 
of the usual needs. They can be used to define the formatting 
of one or more columns.

Each accept a `prefix` and `suffix` argument that can be used 
to add a currency symbol for example. Also they all have 
`na_str` argument (defaut to ""), the string to use when data 
are not available.

- `colformat_num()` with arguments `digits` and `big.mark`
- `colformat_int()` with arguments `big.mark`
- `colformat_char()`

```{r}
dat <- head(mtcars, n = 10)
dat[3:7, 1] <- NA
dat[, 2] <- dat[, 6] * 1000000

ft <- flextable(dat)
num_keys <- c("mpg", "disp", "drat", "wt", "qsec")
int_keys <- c("cyl", "hp", "vs", "am", "gear", "carb")

ft <- colformat_num(x = ft, col_keys = num_keys, big.mark = ",", digits = 2, na_str = "missing")
ft <- colformat_int(x = ft, col_keys = int_keys, big.mark = ",")
autofit(ft)
```


### set_formatter function

It is possible to define the functions that will be used to format the data.frame 
values into strings. `set_formatter` set column formatter functions. 


```{r}
ft <- flextable(head(mtcars, n = 10 ), 
                   col_keys = c("gear", "mpg", "qsec"))
ft <- set_formatter(ft, 
    mpg = function(x) sprintf("%.04f", x),
    gear = function(x) sprintf("%.0f gears", x)
  )
ft <- theme_booktabs(ft)
ft <- autofit(ft)
ft
```


### compose function

`flextable` content can be defined with function `compose`.

It lets user control the formated content at the cell level of the table. 
It is possible to define a content for a row subset and a column as well 
as on the whole column. One can mix images and text (but not with 
PowerPoint because PowerPoint can not do it).

Below the starting point of next illustrations:

```{r}
myft <- flextable( head(mtcars), 
  col_keys = c("am", "separator", "gear", "mpg", "drat", "qsec" ))
myft <- bold(myft, part = "header")
myft <- border(myft, border = fp_border( width = 0), 
  border.top = fp_border(), border.bottom = fp_border(), 
  part = "all")
myft <- align(myft, align = "right", part = "all" )
myft <- border(myft, j = ~ separator, border = fp_border(width=0), part = "all")
myft <- width(myft, j = ~ separator, width = .1)
myft
```

The function requires a call to `as_paragraph` which will concatenate text or images chunks as 
a paragraph. 

![](https://www.ardata.fr/img/illustrations/flextable_compose.svg)


The following example shows how to control the format of displayed values and how to 
associate them with specific text formatting properties (bold red text):

```{r}
myft <- compose( 
  myft, j = "mpg", 
  value = as_paragraph(
    "mpg value is ", 
    as_chunk(sprintf("%.01f", mpg), props = fp_text(color = "red", bold = TRUE) ) )
  )
myft <- autofit(myft)
myft
```

With that system, it's easy to concatenate multiple values:

```{r}
myft <- compose( 
  myft, j = "mpg", 
  value = as_paragraph(
    "mpg value is ", 
    as_chunk(sprintf("%.01f", mpg), props = fp_text(color = "red", bold = TRUE) ), 
    " with ",
    as_chunk(sprintf("# %.0f", carb), props = fp_text(color = "gray", italic = TRUE) )
    )
  )

myft <- autofit(myft)
myft
```

Or to define specific title headers:

```{r}
myft <- compose( 
  myft, j = "mpg", part = "header",
  value = as_paragraph(
    "Miles/(US) gallon ", 
    as_chunk("* with num of carb.", props = fp_text(color = "gray", vertical.align = "superscript") )
    )
  )

myft <- autofit(myft)
myft
```

#### Images

Function `compose` supports images insertion. Use function `as_image` in `as_paragraph` call:

```{r}
img.file <- file.path( R.home("doc"), "html", "logo.jpg" )

myft <- compose( myft, i = ~ qsec > 18, j = "qsec", 
  value = as_paragraph(as_image( src = img.file, width = .20, height = .15))
)
myft <- autofit(myft)
myft
```

#### minibar

Function `compose` supports *mini barplots* insertion. Use function `minibar` in `as_paragraph` call:

```{r}
myft <- flextable( head(iris, n = 10 ))

myft <- compose( myft, j = 1,
  value = as_paragraph(
    minibar(value = Sepal.Length, max = max(Sepal.Length))
  ),
  part = "body")

autofit(myft)
```
