---
title: "Format flextable objects"
author: "David Gohel"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Format flextable objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r warning=FALSE, echo=FALSE}
library(flextable)
library(dplyr)
```

Formatting `flextable` can be made with simple functions. 

## Shortcuts functions

There are simple functions to modify formatting properties of flextable objects: `bg`, `bold`, `border`, `color`, `padding`.

```{r}
myft <- flextable(head(iris))
tabwid(myft)
```

### bold

```{r}
myft <- flextable(head(iris)) %>% 
  # bold header
  bold(part = "header") 
tabwid(myft)
```

### Font size

```{r}
myft <- flextable(head(iris)) %>% 
  # change font size header
  fontsize(part = "header", size = 14) 
tabwid(myft)
```

### change font color

```{r}
myft <- myft %>% color(color = "#E4C994")
tabwid(myft)
```

### Italic

```{r}
myft <- myft %>% 
  italic(i = ~ Sepal.Length > 5, 
         j = ~ Sepal.Length + Sepal.Width, italic = TRUE)
tabwid(myft)
```

### change background color

```{r}
myft <- myft %>% 
  # light gray as background color for header
  bg(bg = "#F0F0F0", part = "header") %>% 
  # dark gray as background color for body
  bg(bg = "#333333", part = "body")

tabwid(myft)
```

### Text alignment

```{r}
myft <- myft %>% align( align = "center", part = "all" )
tabwid(myft)
```


### add padding

```{r}
myft <- myft %>% padding( padding = 3, part = "all" )
tabwid(myft)
```

### change border

```{r}
myft <- myft %>% 
  border( border.top = fp_border(color="orange"), 
          border.bottom = fp_border(color="orange"), 
          border.left = fp_border(color="white"), 
          border.right = fp_border(color="white"), 
          part = "all" )
  
tabwid(myft)
```


## Conditional formatting


Conditional formatting can be made by using the selector arguments.

```{r warning=FALSE, message=FALSE}
myft <- myft %>% 
  color(i = ~ Sepal.Length < 5 & Petal.Length > 1.3, 
        j = ~ Petal.Width + Species, 
        color="red") 
tabwid(myft)
```



## Function style

The `style` function lets you style a selection of the flextable with several formatting properties. 

> It main advantage is to let specify a set of formatting properties for a selection. 

Package officer needs to be loaded, it provides the following formatting properties:

* for text: `fp_text`
* for paragraphs: `fp_par`
* for cells: `fp_cell` and `fp_border`


```{r}
library(officer)
def_cell <- fp_cell(border = fp_border(color="#00C9C9"))
def_par <- fp_par(text.align = "center")
def_text <- fp_text(color="#999999", italic = TRUE)
def_text_header <- update(color="black", def_text, bold = TRUE)

ft <- flextable(head(mtcars, n = 10 )) %>% 
  style( pr_c = def_cell, pr_p = def_par, pr_t = def_text, part = "body")  
tabwid(ft)

ft <- ft %>% 
  style( pr_c = def_cell, pr_p = def_par, pr_t = def_text_header, part = "header")  
tabwid(ft)
```

### Text rotation 

Text rotation is possible in flextable objects. This is achieved by using function `style` and its argument `pr_c`. This need to be a 
`fp_cell` object with `text.direction='btlr'` or `text.direction='tbrl'`.

```{r}
ft <- flextable(head(iris)) %>% 
  style(pr_c = fp_cell(text.direction = "btlr"), 
        part = "header") %>% 
  theme_vanilla() %>% 
  autofit() %>% 
  height(height = 1, part = "header")

tabwid(ft)
```

Note that it only works as expected when `fp_cell` is set 
with `vertical.align = "middle"` which is the default value 
of `fp_cell`. 

## display function

Below the starting point of next illustrations:

```{r}
myft <- flextable(head(mtcars), 
                  col_keys = c("am", "separator", "gear", "mpg", "drat" )) %>% 
  bold(part = "header") %>% 
  border(border = fp_border( width = 0), 
         border.top = fp_border(), 
         border.bottom = fp_border(), part = "all") %>% 
  align(align = "right", part = "all" ) %>%
  border(j = ~ separator, border = fp_border(width=0), part = "all") %>% 
  width(j = ~ separator, width = .1)

tabwid(myft)
```

Flextable content is defined with `display` function. 

It always requires a call to `fpar` which is a simple wrapper whose role is to concatenate  formated text. The following call to fpar only let you modify the display content of a cell (because fpar argument is only a single string). 

```{r}
myft <- myft %>%
  display( 
    mpg = fpar(formatC(mpg, format = "f", digits = 3 ) ) 
  )

tabwid(myft)
```

If more control is needed, i.e. for conditionnal formatting, one can use function `ftext` to associate a text with formatting properties.

```{r}
myft <- myft %>%
  display(i = ~ drat > 3.6, 
    gear = fpar( ftext( gear, prop = fp_text(bold = TRUE, color="red") ) )
  )

tabwid(myft)
```

`fpar` can have multiple arguments that will be then concatenated.

```{r}
myft <- myft %>%
  display(
    mpg = fpar(mpg, " and ", ftext( carb, prop = fp_text(bold = TRUE)), " carb(s)." ) 
  ) %>% autofit()

tabwid(myft)
```

`fpar` handle call to `minibar`. This will generate a simple bar to be inserted as content.

```{r}
myft <- myft %>%
  display( i = ~ gear < 4,
    drat = fpar(minibar( value = drat, max = max(.$drat), 
                         barcol = "#C90000", width = 1, height = .15) )
  ) %>% autofit()

tabwid(myft)
```

Note usage of `.`, this object represents the input dataset of `flextable` call. The expression is not vectorized 





