---
title: "Format `flextable` objects"
author: "David Gohel"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flextable_format}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`flextable` makes possible to create beautiful reporting table from `data.frame`. 

```{r warning=FALSE, message=FALSE}
library(flextable)
library(oxbase)
library(dplyr)

myft <- flextable(head(mtcars)) %>% 
  theme_zebra()
tabwid(myft)
```

## Format table elements

There are functions that are formatting properties of table elements. 

```{r warning=FALSE, message=FALSE}
myft <- myft %>% 
  color(color="darkblue")  %>% 
  color(color="orange", part = "header" )
tabwid(myft)
```

Conditional formatting can be made by using the selector arguments.

```{r warning=FALSE, message=FALSE}
myft <- myft %>% 
  color(i = ~ drat > 3.5 & wt > 2.6, 
        j = ~ vs + am + gear + carb, 
        color="red") 
tabwid(myft)
```

## Shortcuts functions

There are simple functions designed as `color` and that can be used to modify formatting properties of any element of a flextable: `bg`, `bold`, `border`, `color`, `padding`.

```{r}
myft <- flextable(head(iris)) %>% 
  bold(part = "header") %>% 
  padding( padding = 3, part = "all" ) %>% 
  border( border.top = pr_border(color="orange"), 
          border.bottom = pr_border(color="orange"), 
          part = "all" )
  
tabwid(myft)
```

## `style` function

