---
title: "Table layout"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Table layout}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

`flextable` layout can be managed with few functions. 

```{r warning=FALSE, message=FALSE}
library(flextable)
library(officer)
library(magrittr)
```


## Cell merging 

### vertical merging

`merge_v` will merge adjacent duplicated cells for each column of the selection.

```{r warning=FALSE, message=FALSE}
select_columns <- c("Species", "Petal.Length", "Petal.Width")
myft <- flextable(iris[46:55,], col_keys = select_columns) %>% 
  merge_v(~ Species + Petal.Width )

tabwid(myft) 
```

### horizontal merging

`merge_h` will merge adjacent duplicated cells for each row of the selection.

```{r warning=FALSE, message=FALSE}
select_columns <- c("Species", "Petal.Length", "Petal.Width")
myft <- flextable(head(mtcars, n = 10 ) ) %>% 
  merge_h( ) %>%
  border(border = fp_border(), part = "all") # and add borders

tabwid(myft)
```

### general merging function

```{r warning=FALSE, message=FALSE}
select_columns <- c("Species", "Petal.Length", "Petal.Width")
myft <- flextable(head(mtcars, n = 6 ) ) %>% 
  merge_at( i = 1:3, j = 1:3) %>%
  border(border = fp_border(), part = "all") # and add borders

tabwid(myft)
```


### delete merging informations

If you want to get rid of all merging (i.e. for development purposes), use `merge_none`:

```{r}
tabwid(myft %>% merge_none())
```

## Manage headers

### col_keys

Parameter `col_keys` define the variables to display and their order. 

```{r}
library(dplyr)
iris %>% 
  group_by(Species) %>% 
  do( head(., n = 3) ) %>% 
  flextable(col_keys = c("Species", "Sepal.Length", "Petal.Length") ) %>%
  theme_booktabs() %>% 
  tabwid()

```

If parameter `col_keys` has variables that are not existing in the dataset, they will be considered as blank columns and can be used as separators (in fact they can be use as you want, there is only no mapping of data associated).

```{r}
iris %>% group_by(Species) %>% do( head(., n = 3) ) %>%
  flextable(col_keys = c("Species", "col_1", "Sepal.Length", "Petal.Length") ) %>%
  theme_vanilla() %>% autofit() %>% 
  border(j=2, border = fp_border(width=0), part = "all") %>% 
  tabwid()

```

**`col_keys` default values are the names of the data.frame used to fill the flextable.**

### Change labels

Use `set_header_labels` to replace labels of the bottom row of header. 

```{r warning=FALSE, message=FALSE}
ft <- flextable( head( iris ) ) %>% 
  set_header_labels(Sepal.Length = "Sepal", 
    Sepal.Width = "Sepal", Petal.Length = "Petal",
    Petal.Width = "Petal", Species = "Species" )
  
ft %>% theme_vanilla() %>% autofit() %>% tabwid()
```

### add a row of labels

Use `add_header` to add an header row. 

```{r warning=FALSE, message=FALSE}
ft <- ft %>% 
  add_header(Sepal.Length = "length",
    Sepal.Width = "width", Petal.Length = "length",
    Petal.Width = "width", Species = "Species", top = FALSE ) 
ft %>% theme_vanilla() %>% autofit() %>% tabwid()
ft <- ft %>% 
  add_header(Sepal.Length = "Inches",
    Sepal.Width = "Inches", Petal.Length = "Inches",
    Petal.Width = "Inches", Species = "Species", top = TRUE )

# merge identical cells
ft <- ft %>% merge_h(part = "header") %>% merge_v(part = "header")

ft %>% theme_vanilla() %>% autofit() %>% tabwid()
```


### Define headers with a reference table

Use `set_header_df` with a data.frame as parameter. Columns of the dataset will be transposed and joined using a key column. 

#### Input dataset 

Variable `col_keys` define key values to match with flextable column keys (defined by argument `col_keys` of `flextable` function).

This key column will not be displayed. Other variables will added as rows. Note that variables names are not displayed.

```{r warning=FALSE, message=FALSE}
typology <- data.frame(
  col_keys = c( "Sepal.Length", "Sepal.Width", "Petal.Length",
                "Petal.Width", "Species" ),
  type = c("double", "double", "double", "double", "factor"),
  what = c("Sepal", "Sepal", "Petal", "Petal", "Species"),
  measure = c("Length", "Width", "Length", "Width", "Species"),
  stringsAsFactors = FALSE )
typology %>% flextable() %>% theme_vanilla() %>% autofit() %>% tabwid()
```

#### Add the dataset as header rows

Then use `set_header_df` with parameter `key` that specifies name of the column used 
to permform the join. Order of columns matters, first column will be first row, second one will be the second row, etc.

```{r warning=FALSE, message=FALSE}
flextable( head( iris ) ) %>% 
  set_header_df( mapping = typology, key = "col_keys" ) %>% 
  merge_h(part = "header") %>% merge_v(part = "header") %>% 
  theme_vanilla() %>% autofit() %>% tabwid()
```



## Cell widths and heights

The default sizes of flextable columns and rows are calculated with a simple algorithm. This will drive to inadequate rows heights and columns widths in some cases (when data values are wider than headers). You can use function `dim` to get flextable dimensions. 

```{r}
ft_base <- flextable(head(iris)) %>% 
  theme_tron_legacy()
tabwid(ft_base)
dim(ft_base)
```

### Pretty dimensions

Function `dim_pretty` is computing optimized widths and heights. 

```{r}
dim_pretty(ft_base)
```

This function is time consuming and should be used sparingly.

### Adjusts automatically cell widths and heights

Function `autofit` optimises widths and heights of the flextable. This function is almost always 
to be called once when using flextable, it makes compact tables.

```{r}
ft <- ft_base %>% 
  autofit(add_w = 0, add_h = 0)

dim(ft)
tabwid(ft)
```

### Adjusts manually cell widths and heights

Function `width` and `height` let you control dimensions of a flextable. 

```{r}
ft <- ft_base %>% autofit() %>% 
  width(j = ~ Species, width = 2) %>% 
  height(i = 4:5, height = .5)
tabwid(ft)
```


