---
title: "flextable layout"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Table layout}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>", 
  eval = TRUE
)
library(htmltools)
library(magrittr)
library(data.table)

as_ul <- function( names ) {
  do.call( tags$ul, lapply(names, tags$li) )
}

```


`flextable` layout can be easily managed. A set of functions will let you 
merge cells, add title rows, add footer notes, change the withs or heights.

```{r}
library(flextable)
library(officer)
```


## flextable col_keys

Parameter `col_keys` of function `flextable` define the variables to display 
and their order. 

```{r}
data <- iris[c(1:3, 51:53, 101:104),]
myft <- flextable(data, col_keys = c("Species", "Sepal.Length", "Petal.Length") )
myft
```

If parameter `col_keys` has variables that are not existing in the dataset, they 
will be considered as blank columns and can be used as separators (in fact they 
can be use as you want, there is only no mapping of data associated).

```{r}
myft <- flextable(
  data = data, 
  col_keys = c("Species", "col_1", "Sepal.Length", "Petal.Length") )
myft <- theme_vanilla(myft)
myft <- autofit(myft)
myft <- empty_blanks(myft)
myft
```

**`col_keys` default values are the names of the data.frame used to fill the flextable.**

## Change labels

Use `set_header_labels()` to replace labels of the bottom row of header. When the flextable 
is created, their values are the column names of the data.frame. 


```{r}
ft <- flextable( head( iris ) ) 
ft <- set_header_labels(ft, Sepal.Length = "Sepal length", 
    Sepal.Width = "Sepal width", Petal.Length = "Petal length",
    Petal.Width = "Petal width" )
ft
```

New header rows can be added at the top or bottom of the header. 
This part in documented in part 
[Manage headers and footers](#manage-headers-and-footers).

Under the hood, the names are in a single row data.frame 
associated with the header part of the flextable. You can 
add new rows later, they will be binded to that data.frame. 

## Cell merging 

To illustrate functions, we will use a basic flextable example: 

```{r}

dat <- data.frame(
  letters1 = c("a", "b", "b", "c"), 
  letters2 = c("d", "e", "b", "b"), 
  number = 1:4, stringsAsFactors = FALSE )

myft <- flextable(dat)
myft <- theme_box(myft)
myft
```

<div class="row">

<div class="col-12 col-sm-8 col-md-6">

### vertical merging of similar values

`merge_v()` will merge adjacent duplicated cells for each column of the selection.

```{r}
merge_v(myft, j = ~ letters1 + letters2 )
```

</div>

<div class="col-12 col-sm-8 col-md-6">

### horizontal merging of similar values

`merge_h()` will merge adjacent duplicated cells for each row of the selection.

```{r}
merge_h(myft)
```

</div>

</div>
<div class="row">

<div class="col-12 col-sm-8 col-md-6">


### horizontal merging of columns

Function `merge_h_range` is close to the previous one but merge 
all colums between a range of columns.

```{r}
merge_h_range(myft, 
  i =  ~ number < 3, 
  j1 = "letters1", j2 = "letters2")
```


</div>

<div class="col-12 col-sm-8 col-md-6">


### general merging function

`merge_at()` will merge cells for a given continuous selection of rows and cells. 
The result is a single cell.

```{r}
myft %>% 
  merge_at(
    i = 1:2, j = 1:2)
```

</div>

</div>



### delete merging informations

If you want to get rid of all merging (i.e. for development purposes), use `merge_none()`:

```{r}
merge_none(myft)
```


### Borders and merging

When cells are merged, the rendered borders will be those of the first cell. If a 
column is made of three merged cells, the bottom border that will be seen will be 
the bottom border of the first cell in the column. From a user point of view, this 
is wrong, the bottom should be the one defined for cell 3. Function `fix_border_issues` 
is trying to fix that issue. 

```{r}
ft <- data.frame(a = 1:5, b = 6:10) %>%
  flextable() %>% theme_box() %>%
  merge_at(i = 4:5, j = 1, part = "body") %>%
  hline(i = 5, part = "body",
        border = fp_border(color = "orange", width = 3) )
ft
fix_border_issues(ft)
```



## Manage headers and footers

Use one of the following functions to add an header row or a footer row:

```{r echo=FALSE}
as_ul(ls(envir = as.environment("package:flextable"), pattern = "^add_(header|footer)"))
```

We will mainly demonstrate headers but same can be apply with footers.

```{r}
ft <- flextable( head( iris ) ) 
ft <- set_header_labels(ft, Sepal.Length = "Sepal", 
    Sepal.Width = "Sepal", Petal.Length = "Petal",
    Petal.Width = "Petal" )
ft
```


### add a row of labels

<div class="row">

<div class="col-12 col-sm-8 col-md-6">


```{r}
ft <- add_header(ft, 
  Sepal.Length = "length", 
  Sepal.Width = "width", 
  Petal.Length = "length", 
  Petal.Width = "width", 
  top = FALSE )
ft <- theme_booktabs(ft)
```

</div>

<div class="col-12 col-sm-8 col-md-6">

```{r echo=FALSE}
ft
```

</div>

</div>


### lines of text

<div class="row">

<div class="col-12 col-sm-8 col-md-6">

```{r}
ft <- add_header_lines(ft, 
  values = c("this is a first line", 
     "this is a second line") ) 
ft <- theme_booktabs(ft)
```

</div>

<div class="col-12 col-sm-8 col-md-6">

```{r echo=FALSE}
ft
```

</div>

</div>




### Define headers with a reference table

Use `set_header_df()` with a data.frame as parameter. Columns of 
the dataset will be transposed and joined using a key column. 

1. The reference table 

Variable `col_keys` define key values to match with flextable 
column keys (defined by argument `col_keys` of `flextable()` function).

This key column will not be displayed. Other variables will added as 
rows. Note that variables names are not displayed. 

<div class="row">

<div class="col-12 col-sm-8 col-md-6">

```{r}
typology <- data.frame(
  col_keys = c( "Sepal.Length", 
    "Sepal.Width", "Petal.Length",
    "Petal.Width", "Species" ),
  type = c("double", "double", "double", 
    "double", "factor"),
  what = c("Sepal", "Sepal", "Petal", 
    "Petal", "Species"),
  measure = c("Length", "Width", "Length", 
    "Width", "Species"),
  stringsAsFactors = FALSE )
```

</div>

<div class="col-12 col-sm-8 col-md-6">

```{r echo=FALSE}
autofit( theme_vanilla(flextable(typology)) )
```

</div>

</div>


2. Use it as header rows

Then use `set_header_df()` with parameter `key`.  `key` is the name of the column used 
to permform the join operation. 

> Order of columns matters, first column will be first row, second one will 
  be the second row, etc.

<div class="row">

<div class="col-12 col-sm-8 col-md-6">

```{r}
ft <- flextable( head( iris ) )
ft <- set_header_df( ft, mapping = typology, key = "col_keys" )

ft <- merge_h(ft, part = "header")
ft <- merge_v(ft, part = "header")

ft <- theme_booktabs(ft)
ft <- autofit(ft)
ft <- fix_border_issues(ft)
```


</div>

<div class="col-12 col-sm-8 col-md-6">

```{r echo=FALSE}
ft
```

</div>

</div>




## Cell widths and heights

The default sizes of flextable columns and rows are set by default values. 
This will drive to inadequate rows heights and columns widths in some cases. 
You can use function `dim` to get flextable dimensions. 

```{r}
ft_base <- flextable(head(mtcars))
ft_base <- theme_vader(ft_base, fontsize = 13)
ft_base
dim(ft_base)
```

### Pretty dimensions

Function `dim_pretty()` is computing optimized widths and heights. 

```{r}
dim_pretty(ft_base)
```

### Adjusts automatically cell widths and heights

Function `autofit()` optimises widths and heights of the flextable. 
This function is almost always to be called once when using 
flextable objects, it makes compact tables.

```{r}
ft <- autofit(ft_base, add_w = 0, add_h = 0)

dim(ft)
ft
```

> Soft returns (*a line break in a paragraph*) support : function `autofit` and `dim_pretty` do not
support soft returns and may return wrong results (\n will be considered as "").
  
### Adjusts manually cell widths and heights

Function `width()` and `height()` let you control dimensions of a flextable. `height_all()` is an helper function to set the same height to each part of the table.

```{r}
ft <- autofit(ft_base)
ft <- width(ft, j = ~ mpg + cyl + disp, width = 2)
ft <- height_all( ft, height = .4 )
ft <- height( ft, i = 3, height = 1 )
ft
```

## Groups as row titles


Package flextable does not support data transformation. A *grouped* dataset 
is then needed (and *tidy*). This has to be done prior to the creation of the object. 

Function `as_grouped_data` will modify data structure so that it will be 
easy to manage grouped data representation. Repeated consecutive values 
of group columns will be used to define the title of the groups and will 
be added as a *row title*.

Let's have an example with aggragated data from dataset `CO2`:

```{r}
library(data.table)
data_CO2 <- dcast(as.data.table(CO2), 
  Treatment + conc ~ Type, value.var = "uptake", fun.aggregate = mean)
head(data_CO2)
```

`as_grouped_data` will restructure the dataset:

```{r}
data_CO2 <- as_grouped_data(x = data_CO2, groups = c("Treatment"))
head(data_CO2)
```

The result is suitable for method `as_flextable`. A call to this function 
and few formatting operations are producing the following result:

```{r}
zz <- as_flextable( data_CO2 ) %>% 
  bold(j = 1, i = ~ !is.na(Treatment), bold = TRUE, part = "body" ) %>% 
  bold(part = "header", bold = TRUE ) %>% 
  width(width = 1.5)
zz
```

It's now easier to customize the rendering. Let's format column `conc` 
as an integer column:

```{r}
zz <- zz %>% 
  compose(i = ~ !is.na(conc), j = "conc", 
          value = as_paragraph(
            as_chunk(conc, formater = function(x) sprintf("%.0f", x))
          )
  )
zz
```

Now let's add nice bars before displaying the figures:

```{r}
zz <- zz %>% 
  compose(i = ~ is.na(Treatment), j = "Quebec", 
          value = as_paragraph(
            minibar(Quebec), 
            " ", 
            as_chunk(Quebec, formater = function(x) sprintf("%.01f", x))
            )
          ) %>% 
  compose(i = ~ is.na(Treatment), j = "Mississippi", 
          value = as_paragraph( minibar(Mississippi), 
                                " ",
                                as_chunk(Mississippi, 
                                         formater = function(x) sprintf("%.01f", x) )
                                )
          ) %>% 
  align(j = 2:3, align = "left")
zz
```

And finally, add a footnote in the footer part:

```{r}
add_footer_lines(zz, "dataset CO2 has been used for this flextable") 
```


## xtable 

`xtable` objects can be transformed as flextable objects with function `xtable_to_flextable()`. 

```{r}
if( require("xtable") ){
  temp.ts <- ts(cumsum(1 + round(rnorm(100), 0)),
    start = c(1954, 7), frequency = 12)
  ft <- xtable_to_flextable(x = xtable(temp.ts, digits = 0),
    NA.string = "-")
  ft
}
```

