---
title: "flextable overview"
author: "David Gohel"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{flextable overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction 

`flextable` can easily create reporting table from `data.frame`. You can merge cells, add header rows, change any format and specify how data should be displayed in cells.

```{r warning=FALSE}
library(flextable)
library(officer)
library(dplyr)

typology <- data.frame(
  col_keys = c( "Sepal.Length", "Sepal.Width", "Petal.Length",
                "Petal.Width", "Species" ),
  what = c("Sepal", "Sepal", "Petal", "Petal", "Species"),
  measure = c("Length", "Width", "Length", "Width", "Species"),
  stringsAsFactors = FALSE )
typology

data <- iris %>% 
  group_by(Species) %>% 
  do( head(., n = 3) )
data

flextable(data) %>% 
  set_header_df(mapping = typology, key = "col_keys" ) %>% 
  theme_vanilla() %>% autofit() %>% 
  tabwid()
```

---


## Getting started

Let's have a step by step demo. First create a flextable and change header font in bold. Function `tabwid` will wrap it in an *htmlwidget*.

```{r warning=FALSE, message=FALSE}
library(flextable)
library(officer)
library(dplyr)

myft <- flextable(head(mtcars), 
                  col_keys = c("am", "carb", "gear", "mpg", "drat" ))
tabwid(myft)
```

**flextable function**: `flextable` create a flextable object based on input data. `flextable` does not format automatically the table but it is preparing the object to be formatted. Optional argument `col_keys` is used to only display a subset of columns. 

**tabwid function**: `tabwid` is the function that transform that *flextable* object into an html widget object (in r markdown documents or shiny applications). We need it here as we are working in an r markdown document.

Let's keep it simple and apply a theme to format the whole table. Functions `theme_` are sugar functions whose role is to apply a set of formatting instructions to a flextable. For example, `theme_vanilla` set specific borders, right align paragraphs and make headers bold. 

```{r warning=FALSE, message=FALSE}
myft <- myft %>% theme_vanilla()
tabwid(myft)
```

### Layout

Table layout can be modified. Man can add or change header rows, change cells height and width and merge cells. Also, there is an important function named `autofit`.

We will use `merge_v` to merge identical consecutive cells of 
columns "carb" and "am".

```{r warning=FALSE, message=FALSE}
myft <- myft %>%
  merge_v(j = c("am", "carb") )
tabwid(myft)
```

`set_header_labels` set labels:

```{r warning=FALSE, message=FALSE}
myft <- myft %>%
  set_header_labels( carb = "number of carb." )
tabwid(myft)
```



### Formating

Many sugar functions can be used to format flextables: `bg`, `fontsize`, `italic`, `bold`, `color`, `padding`, `border`. 

```{r}
myft <- myft %>% italic(j = 1) %>% 
  bg(bg = "#CCDDCCCC", part = "header")
  
tabwid(myft)
```

When no sugar function is available, use `style`. It modify cells, paragraphs or text properties. 

```{r}
myft <- myft %>% 
  style(pr_p = fp_par(text.align = "center", padding = 3), 
        pr_c = fp_cell(border=fp_border(color = "#BEBFBB", width = 2)),
        pr_t = fp_text(font.size = 11, color = "#000011"),
        part = "all" ) %>% 
  bold(part = "header") %>% fontsize(size = 12, part = "header") %>% 
  autofit()

tabwid(myft)
```

`autofit` adjust widths and heights of cells. This is the last operation as some operations make columns wider, e.g. changing font size, changing font weight. `autofit` makes sure that any content is displayed as a single line of text.

### Conditionnal formating

Conditional formatting can be made by using the selector arguments. All formatting functions are accepting selector arguments. 

```{r warning=FALSE, message=FALSE}
myft <- myft %>% 
  color(~ drat > 3.5, ~ drat, color = "red") %>% 
  italic(~ drat > 3.5, ~ drat, italic = TRUE) 

tabwid(myft)
```

## Outputs

Flextables can be inserted in r markdown documents and in shiny applications. Use `tabwid` in these cases, when working in RStudio, flextable will be printed in the rstudio viewer pane.

To add these objects in PowerPoint or Word documents, use functions `ph_with_flextable` or `body_add_flextable`. 

```{r}
library(officer)
ft <- flextable(head(mtcars)) %>% 
  theme_zebra() %>% 
  autofit()

read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_flextable(value = ft, type = "body") %>% 
  print(target = "test.pptx")

read_docx() %>% 
  body_add_flextable(value = ft) %>% 
  print(target = "test.docx")
```

