---
title: "flextable overview"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{flextable overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message=FALSE, warning=FALSE}
dir.create("assets/docx", recursive = TRUE, showWarnings = FALSE)
dir.create("assets/pptx", recursive = TRUE, showWarnings = FALSE)
office_doc_link <- function(url){
  stopifnot(requireNamespace("htmltools", quietly = TRUE))
  htmltools::tags$p(  htmltools::tags$span("Download file "),
    htmltools::tags$a(basename(url), href = url), 
    htmltools::tags$span(" - view with"),
    htmltools::tags$a("office web viewer", target="_blank", 
      href = paste0("https://view.officeapps.live.com/op/view.aspx?src=", url)
      ), 
    style="text-align:center;font-style:italic;color:gray;"
    )
}
```

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction 

`flextable` can easily create reporting table from `data.frame`. You can merge cells, add header rows, change any format and specify how data should be displayed in cells. `flextable` objects can be rendered in HTML format but also in Microsoft Word and PowerPoint documents.

Two functions are provided: `flextable` and `regulartable`.

`flextable` function is producing flexible tables where each cell can contain 
several chunks of text with their own set of formatting properties 
(bold, font color, etc.). Function `display` lets customise text of cells (See 
[display function](format.html#display-function)).


`regulartable` function has been written because the first one is ressource 
consumming. The only difference is that it is producing tables where cells 
can contain only one chunk of text with its own set of formatting properties.
Function `set_formatter` let customise text of cells.

> The following table is made with `regulartable`.

```{r warning=FALSE, echo=FALSE}
library(flextable)
library(officer)
library(dplyr)

typology <- data.frame(
  col_keys = c( "Sepal.Length", "Sepal.Width", "Petal.Length",
                "Petal.Width", "Species" ),
  what = c("Sepal", "Sepal", "Petal", "Petal", "Species"),
  measure = c("Length", "Width", "Length", "Width", "Species"),
  stringsAsFactors = FALSE )


data <- iris %>% 
  group_by(Species) %>% 
  do( head(., n = 3) )

ft <- regulartable(data, 
          col_keys = c("Species", "sep_1", "Sepal.Length", "Sepal.Width", "sep_2",  "Petal.Length", "Petal.Width" ) ) %>% 
  set_header_df(mapping = typology, key = "col_keys" ) %>% 
  merge_h(part = "header") %>% 
  merge_v(j = "Species", part = "body") %>% 
  merge_v(j = "Species", part = "header") %>% 
  theme_vanilla() %>% empty_blanks() %>% autofit() 

tabwid(ft)
```

---


## Getting started

Let's have a step by step demo. First create a regulartable and change header font in bold. Function `tabwid` will wrap it in an *htmlwidget*.

```{r warning=FALSE, message=FALSE}
library(flextable)
library(officer)
library(dplyr)

myft <- regulartable(head(mtcars), 
                  col_keys = c("am", "carb", "gear", "mpg", "drat" ))
tabwid(myft)
```

**flextable or regulartable function**: `regulartable` and `flextable` create a flextable object based on input data. Optional argument `col_keys` is used to only display a subset of columns. 

**tabwid function**: `tabwid` is the function that transform that *flextable* object into an html widget object (in r markdown documents or shiny applications). We need it here as we are working in an rmarkdown document.

Let's keep it simple and apply a theme to format the whole table. Functions `theme_` are sugar functions whose role is to apply a set of formatting instructions to a flextable. For example, `theme_vanilla` set specific borders, right align paragraphs and make headers bold. 

```{r warning=FALSE, message=FALSE}
myft <- myft %>% theme_vanilla()
tabwid(myft)
```

### Layout

> To learn more, see article about [layouts](./layout.html).

Table layout can be modified. One can add or change header rows, change cells height and width and merge cells. Also, there is an important function named `autofit` (it adjusts widths and heights regarding to text widths and heights).

We will use `merge_v` to merge identical consecutive cells of 
columns "carb" and "am".

```{r warning=FALSE, message=FALSE}
myft <- myft %>%
  merge_v(j = c("am", "carb") )
tabwid(myft)
```

Cells can be merged with functions `merge_none`, `merge_v` and `merge_h`.

`set_header_labels` set labels:

```{r warning=FALSE, message=FALSE}
myft <- myft %>%
  set_header_labels( carb = "# carb." ) %>% 
  width(width = .75) # set width of all columns to .75 in
tabwid(myft)
```

Headers can be modified with functions `set_header_df`, `set_header_labels` and `add_header`.

`autofit` adjust widths and heights of cells. This should be the last operation as some operations make columns wider, e.g. changing font size, changing font weight. **`autofit` makes sure that any content is displayed as a single line of text.**

```{r warning=FALSE, message=FALSE}
myft <- myft %>% autofit()

tabwid(myft)
```


### Formating

> To learn more, see article about [format](./format.html).

Many sugar functions can be used to format flextables: `bg`, `fontsize`, `italic`, `bold`, `color`, `padding`, `border`. 

```{r}
myft <- myft %>% italic(j = 1) %>% 
  bg(bg = "#C90000", part = "header") %>% 
  color(color = "white", part = "header") %>% 
  border(border = fp_border(color = "orange"), part = "all")
  
tabwid(myft)
```

Conditional formatting can be made by using the selector arguments. All formatting functions are accepting selector arguments. 

```{r warning=FALSE, message=FALSE}
myft <- myft %>% 
  color(~ drat > 3.5, ~ drat, color = "red") %>% 
  bold(~ drat > 3.5, ~ drat, bold = TRUE) %>% 
  autofit()

tabwid(myft)
```


## Outputs

### HTML

Flextables can be inserted in r markdown documents and in shiny applications; in these cases, use `tabwid`. When working in RStudio, flextable will be printed in the rstudio viewer pane and call to `tabwid` is not necessary. Note that flextables are not designed to work with flexdashboard documents.

### PowerPoint or Word documents

To add these objects in PowerPoint or Word documents, use functions:
- `ph_with_flextable` or `ph_with_flextable_at` (PowerPoint)
- `body_add_flextable` (Word)

**`officer` package is required to create a PowerPoint or Word document.**

```{r}
library(officer)
```

The following is producing a PowerPoint document:

```{r}
ft <- regulartable(head(mtcars)) %>% 
  theme_booktabs() %>% 
  autofit()

ppt <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_flextable(value = ft, type = "body") 

print(ppt, target = "assets/pptx/example.pptx") %>% invisible()
```

```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/flextable/articles/", "assets/pptx/example.pptx" ) )
```

The following is producing a Word document:

```{r}
doc <- read_docx() %>% 
  body_add_flextable(value = ft)
print(doc, target = "assets/docx/example.docx") %>% invisible()
```

```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/flextable/articles/", "assets/docx/example.docx" ) )
```
