---
title: "flextable overview"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{flextable overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(flextable)
library(magrittr)

dir.create("assets/docx", recursive = TRUE, showWarnings = FALSE)
dir.create("assets/pptx", recursive = TRUE, showWarnings = FALSE)
office_doc_link <- function(url){
  stopifnot(requireNamespace("htmltools", quietly = TRUE))
  htmltools::tags$p(  htmltools::tags$span("Download file "),
    htmltools::tags$a(basename(url), href = url), 
    htmltools::tags$span(" - view with"),
    htmltools::tags$a("office web viewer", target="_blank", 
      href = paste0("https://view.officeapps.live.com/op/view.aspx?src=", url)
      ), 
    style="text-align:center;font-style:italic;color:gray;"
    )
}

knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>", 
  eval = !is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))
)
```

## Introduction 

`flextable` can easily create reporting table from `data.frame`. You 
can merge cells, add header rows, add footer rows, change any format 
and specify how data should be displayed in cells. `flextable` objects 
can be rendered in HTML format but also in Microsoft Word and 
PowerPoint documents.

> The following table is made with function `flextable` and dataset data:

```{r}
data <- iris[c(1:3, 51:53, 101:104),]
```


```{r warning=FALSE, echo=FALSE}
library(flextable)
library(officer)

typology <- data.frame(
  col_keys = c( "Sepal.Length", "Sepal.Width", "Petal.Length",
                "Petal.Width", "Species" ),
  what = c("Sepal", "Sepal", "Petal", "Petal", "Species"),
  measure = c("Length", "Width", "Length", "Width", "Species"),
  stringsAsFactors = FALSE )

ft <- flextable(
  data, 
  col_keys = c("Species", "sep_1", "Sepal.Length", "Sepal.Width", 
               "sep_2",  "Petal.Length", "Petal.Width" ) )

ft <- set_header_df(ft, mapping = typology, key = "col_keys" )
ft <- merge_h(ft, part = "header")
ft <- merge_v(ft, j = "Species", part = "body")
ft <- merge_v(ft, j = "Species", part = "header")
ft <- theme_booktabs(ft)
ft <- empty_blanks(ft)
ft <- fix_border_issues(ft)
autofit(ft) 
```

## Supported output formats

When working in RStudio, flextable will be printed in the rstudio viewer pane. 

### R Markdown documents

One great feature of `flextables` is that it can be used in R Markdown 
documents; HTML, Word and PowerPoint outputs are managed with 
method `knitr::knit_print`. 

Print the `flextable` object in the rmarkdown's \code{render} chunk 
as demonstrated in this document (...generated from an R Markdown 
document).

````

> this is how to print a flextable in a R Markdown document

```{r eval=FALSE}
library(magrittr)
mytable <- cars %>% 
  head() %>% 
  flextable() %>% 
  autofit()
  
mytable
```

````

#### pandoc versions

Supported formats require some minimum [pandoc](https://pandoc.org/installing.html) versions: 

|    R Markdown output | pandoc version |                 comment                |
|---------------------:|:--------------:|:--------------------------------------:|
|                 HTML |     >= 1.12    |                                        |
|       Microsoft Word |     >= 2.0     | require a recent RStudio (> june 2018) |
| Microsoft PowerPoint |     >= 2.4     |       require pandoc installation      |


### PowerPoint or Word documents with package officer

To add these objects in PowerPoint or Word documents, use functions:

- `ph_with_flextable()` or `ph_with_flextable_at()` (PowerPoint)
- `body_add_flextable()` (Word)

**`officer` package is required to create a PowerPoint or Word document.**

```{r}
library(officer)
```

The following is producing a PowerPoint document:

```{r results='hide'}
ft <- flextable(head(mtcars))
ft <- theme_booktabs(ft)
ft <- autofit(ft)

ppt <- read_pptx()
ppt <- add_slide(ppt, layout = "Title and Content", master = "Office Theme")
ppt <- ph_with_flextable(ppt, value = ft, type = "body") 

print(ppt, target = "assets/pptx/example.pptx")
```

```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/flextable/articles/", "assets/pptx/example.pptx" ) )
```

The following is producing a Word document:

```{r results='hide'}
doc <- read_docx()
doc <- body_add_flextable(doc, value = ft)
print(doc, target = "assets/docx/example.docx")
```

```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/flextable/articles/", "assets/docx/example.docx" ) )
```



---


## Getting started

The main function is `flextable`.

`flextable` function is producing flexible tables where each cell can contain 
several chunks of text with their own set of formatting properties 
(bold, font color, etc.). 

The function is using default values to format the table. It is also using 
a simple formatting function to create the paragraphs of text that will 
be rendered in each cell. All these default values can be modified after 
the flextable creation. 

The following illustration is presenting most of the functions 
available to format a flextable:

![](https://www.ardata.fr/img/illustrations/flextable_functions.svg)

### flextable keypoints

* A flextable is an object that will produce a reporting table from a 
`data.frame` object. 

* A flextable is containing three parts, header, body and footer. 
By default, there is only one header row containing the names of the data.frame. 
The body part is containing data from the data.frame. The footer 
part is not implemented by default.

```{r echo=FALSE}
dmodel <- data.frame( 
  header = rep("body", 4 ),
  header_1 = c("data[1,1]", "data[2,1]", "...", "data[n,1]" ),
  header_2 = c("data[1,2]", "data[2,2]", "...", "data[n,2]" ),
  header_3 = c("data[1,3]", "data[2,3]", "...", "data[n,3]" ),
  header_4 = c("data[1,4]", "data[2,4]", "...", "data[n,4]" ),
  header_5 = c("data[1,5]", "data[2,5]", "...", "data[n,5]" ), 
  stringsAsFactors = FALSE
)
header_model <- data.frame(
  title1 = c( "header", rep("row 1 - cell from 1 to 5", 5) ),
  title2 = c( "header", rep("row 2 - cell from 1 to 2", 2), rep("row 2 cell from 3 to 5", 3) ),
  title3 = c( "header", sprintf("row 3 cell %.0f", 1:5) ), 
  key = c( names(dmodel)), 
  stringsAsFactors = FALSE )

data_model <- flextable( data = dmodel, 
                            col_keys = c("header", "blank", "header_1", "header_2", "header_3", "header_4", "header_5" ) )
data_model <- set_header_df(data_model, mapping = header_model, key = "key")
data_model <- merge_h(data_model, part = "header")

data_model <- add_footer(data_model, 
                         header = "footer",
                         header_1 = "Grouped footer 1.1", 
                         header_2 = "Grouped footer 1.1", 
                         header_3 = "Grouped footer 1.2", 
                         header_4 = "Grouped footer 1.2", 
                         header_5 = "Grouped footer 1.2" )
data_model <- add_footer(data_model, top = FALSE,
                         header = "footer",
                         header_1 = "Grouped footer 2.1", 
                         header_2 = "Grouped footer 2.1", 
                         header_3 = "Grouped footer 2.1", 
                         header_4 = "Grouped footer 2.1", 
                         header_5 = "Grouped footer 2.1" )
data_model <- merge_h(data_model, part = "footer")
data_model <- merge_v(data_model, part = "footer", j = 1)
data_model <- merge_v(data_model, part = "body", j = 1)
data_model <- merge_v(data_model, part = "header", j = 1)
data_model <- theme_box(data_model)
data_model <- flextable::align(data_model, align = "center", part = "all")
data_model <- bg(data_model, bg = "#eb5555", part = "header")
data_model <- bg(data_model, bg = "#475F77", part = "body")
data_model <- bg(data_model, bg = "#2BBBD8", part = "footer")
data_model <- color(data_model, color = "white", part = "footer")
# data_model <- bg(data_model, j = 1, bg = "#475F77", part = "all")
data_model <- width(data_model, j = 1, width = 1.2)
data_model <- width(data_model, j = 2, width = .2)
data_model <- width(data_model, j = 3:7, width = 1.2)

std_border <- fp_border(width = 1.5, color = "white")

data_model <- border_outer(data_model, part="all", border = std_border )
data_model <- border_inner_h(data_model, border = std_border, part="all")
data_model <- border_inner_v(data_model, border = std_border, part="all")

data_model <- empty_blanks(data_model)
data_model <- padding(data_model, padding = 4, part = "all")
data_model <- fontsize(data_model, size = 12, part = "all")
data_model <- italic(data_model, italic = FALSE, part = "footer")
data_model <- bold(data_model, bold = TRUE, part = "all")
data_model <- color(data_model, color = "white", part = "all")
data_model
```


* Any formatting property related to text, paragraphs, 
cells in the table can be modified. 


* A cell is made of one single paragraph of text. Paragraphs 
can contain several chunks of text with different formatting 
but also images. 

* The dimensions of the flextable elements, widths of columns and 
heights of rows, are always defined. Beeing able to define the exact 
size necessary to **display data on a single line** is a key concept. 
See functions `autofit()` and `dim_pretty()`. 

* Columns to display (`col_keys`) are by default all columns 
of the data.frame but can be choosen from a subset of existing 
columns and also unexisting columns. Unexisting columns are 
containing blanks by default but this content can be customized.


Let's have a step by step demo. First create a flextable and change header font in bold. 

```{r warning=FALSE}
library(flextable)
library(officer)

myft <- flextable(
  head(mtcars), 
  col_keys = c("am", "carb", "gear", "mpg", "drat" ))
myft
```

**flextable**: it is creating a flextable object based on input data. 
Optional argument `col_keys` is used to only display a subset of columns. 

Functions `theme_` are sugar functions whose role is to apply a set 
of formatting instructions to a flextable. For example, `theme_vanilla` 
set specific borders, right align paragraphs and make headers bold. 

```{r warning=FALSE}
myft <- theme_vanilla(myft)
myft
```

### Layout

Table layout can be modified. One can add or change header/footer rows, change cells height and width and merge cells. 

Also, there is an important function named `autofit` (it adjusts widths and heights regarding to text widths and heights).


```{r warning=FALSE}
myft <- merge_v(myft, j = c("am", "carb") )
myft <- set_header_labels( myft, carb = "# carb." )
myft <- autofit(myft)
myft
```

To learn more, see article about [layouts](./layout.html).


### Formatting

> To learn more, see article about [format](./format.html).

Many sugar functions can be used to format flextables: `bg()`, `fontsize()`, `italic()`, `bold()`, `color()`, `padding()`... 

Conditional formatting can be made by using the selector arguments. All formatting functions are accepting selector arguments. 

```{r}
myft <- italic(myft, j = 1)
myft <- bg(myft, bg = "#C90000", part = "header")
myft <- color(myft, color = "white", part = "header")
myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)
myft <- autofit(myft)
myft
```

## Selectors

Selectors can be used to specify the rows and columns where the operation should happened.

Many `flextable` functions have *selectors* `i` and `j`: `bg`, `bold`, 
`border`, `color`, `padding`, `fontsize`, `italic`, `align`, ... 

`i` for rows selection and `j` for columns selection can be expressed in different ways:

* as a formula: `i = ~ Species %in% "versicolor", j = ~ Species + Sepal.Length`

```{r}
ft <- flextable(head(iris, n = 10))
ft <- color(ft, i = ~ Sepal.Length < 5, j = ~ Sepal.Length + Sepal.Width, color = "orange")
ft
```

* as an integer vector, each element is the row number or `col_key` number

```{r}
ft <- flextable(head(iris, n = 10))
ft <- color(ft, i = 1:3, j = 1:3, color = "orange")
ft
```

* as a logical vector; for `j`, a p length if p columns; for `i`, a n length if n rows

```{r}
dat <- head(iris, n = 10)
ft <- flextable(dat)

ft <- color(ft, i = dat$Sepal.Length < 5, color = "orange")
ft
```

* j support also simple character vector containing the `col_key` names.

```{r}
dat <- head(iris, n = 10)
ft <- flextable(dat)
ft <- color(ft, j = "Sepal.Length", color = "orange", part = "all")
ft
```



