---
title: "flextable overview"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{flextable overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction 

`flextable` can easily create reporting table from `data.frame`. You can merge cells, add header rows, change any format and specify how data should be displayed in cells. `flextable` objects can be rendered in HTML format but also in Microsoft Word and PowerPoint documents.

The following table is made with flextable.

```{r warning=FALSE, echo=FALSE}
library(flextable)
library(officer)
library(dplyr)

typology <- data.frame(
  col_keys = c( "Sepal.Length", "Sepal.Width", "Petal.Length",
                "Petal.Width", "Species" ),
  what = c("Sepal", "Sepal", "Petal", "Petal", "Species"),
  measure = c("Length", "Width", "Length", "Width", "Species"),
  stringsAsFactors = FALSE )


data <- iris %>% 
  group_by(Species) %>% 
  do( head(., n = 3) )

ft <- flextable(data, 
          col_keys = c("Species", "sep_1", "Sepal.Length", "Sepal.Width", "sep_2",  "Petal.Length", "Petal.Width" ) ) %>% 
  set_header_df(mapping = typology, key = "col_keys" ) %>% 
  merge_v(j = "Species") %>% 
  theme_vanilla() %>% empty_blanks() %>% autofit() 

tabwid(ft)
```

The input dataset and the metadata for columns labels are printed below:

```{r}
data
typology
```


---


## Getting started

Let's have a step by step demo. First create a flextable and change header font in bold. Function `tabwid` will wrap it in an *htmlwidget*.

```{r warning=FALSE, message=FALSE}
library(flextable)
library(officer)
library(dplyr)

myft <- flextable(head(mtcars), 
                  col_keys = c("am", "carb", "gear", "mpg", "drat" ))
tabwid(myft)
```

**flextable function**: `flextable` create a flextable object based on input data. `flextable` does not format automatically the table but it is preparing the object to be formatted. Optional argument `col_keys` is used to only display a subset of columns. 

**tabwid function**: `tabwid` is the function that transform that *flextable* object into an html widget object (in r markdown documents or shiny applications). We need it here as we are working in an r markdown document.

Let's keep it simple and apply a theme to format the whole table. Functions `theme_` are sugar functions whose role is to apply a set of formatting instructions to a flextable. For example, `theme_vanilla` set specific borders, right align paragraphs and make headers bold. 

```{r warning=FALSE, message=FALSE}
myft <- myft %>% theme_vanilla()
tabwid(myft)
```

### Layout

Table layout can be modified. Man can add or change header rows, change cells height and width and merge cells. Also, there is an important function named `autofit`.

We will use `merge_v` to merge identical consecutive cells of 
columns "carb" and "am".

```{r warning=FALSE, message=FALSE}
myft <- myft %>%
  merge_v(j = c("am", "carb") )
tabwid(myft)
```

Cells can be merged with functions `merge_none`, `merge_v` and `merge_h`.

`set_header_labels` set labels:

```{r warning=FALSE, message=FALSE}
myft <- myft %>%
  set_header_labels( carb = "# carb." )
tabwid(myft)
```

Headers can be modified with functions `set_header_df`, `set_header_labels` and `add_header`.


### Formating

Many sugar functions can be used to format flextables: `bg`, `fontsize`, `italic`, `bold`, `color`, `padding`, `border`. 

```{r}
myft <- myft %>% italic(j = 1) %>% 
  bg(bg = "#C90000", part = "header") %>% 
  color(color = "white", part = "header") %>% 
  border(border = fp_border(color = "orange"), part = "all")
  
tabwid(myft)
```

Conditional formatting can be made by using the selector arguments. All formatting functions are accepting selector arguments. 

```{r warning=FALSE, message=FALSE}
myft <- myft %>% 
  color(~ drat > 3.5, ~ drat, color = "red") %>% 
  bold(~ drat > 3.5, ~ drat, bold = TRUE) 

tabwid(myft)
```


`autofit` adjust widths and heights of cells. This is the last operation as some operations make columns wider, e.g. changing font size, changing font weight. `autofit` makes sure that any content is displayed as a single line of text.

```{r warning=FALSE, message=FALSE}
myft <- myft %>% autofit()

tabwid(myft)
```


## Outputs

Flextables can be inserted in r markdown documents and in shiny applications. Use `tabwid` in these cases, when working in RStudio, flextable will be printed in the rstudio viewer pane (call to `tabwid` is not necessary in this case). Note that flextables are not designed to work with flexdashboard documents.

To add these objects in PowerPoint or Word documents, use functions:
- `ph_with_flextable` (PowerPoint)
- `body_add_flextable` (Word)

`officer` package is required to create a PowerPoint or Word document.

```{r}
library(officer)
ft <- flextable(head(mtcars)) %>% 
  theme_zebra() %>% 
  autofit()

ppt <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_flextable(value = ft, type = "body") 
if( interactive() ) print(ppt, target = "test.pptx")

doc <- read_docx() %>% 
  body_add_flextable(value = ft)
if( interactive() ) print(doc, target = "test.docx")
```

