---
title: "flextable overview"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{flextable overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(flextable)
library(magrittr)
library(htmltools)
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>", 
  eval = !is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))
)
```

## Selectors

Selectors can be used to specify the rows and columns where the operation should happened.

Many `flextable` functions have *selectors* `i` and `j`: `bg`, `bold`, 
`border`, `color`, `padding`, `fontsize`, `italic`, `align`, ... 

`i` for rows selection and `j` for columns selection can be expressed in different ways:

* as a formula: `i = ~ Species %in% "versicolor", j = ~ Species + Sepal.Length`

```{r}
ft <- flextable(head(iris, n = 10))
ft <- color(ft, i = ~ Sepal.Length < 5, j = ~ Sepal.Length + Sepal.Width, color = "orange")
ft
```

* as an integer vector, each element is the row number or `col_key` number

```{r}
ft <- flextable(head(iris, n = 10))
ft <- color(ft, i = 1:3, j = 1:3, color = "orange")
ft
```

* as a logical vector; for `j`, a p length if p columns; for `i`, a n length if n rows

```{r}
dat <- head(iris, n = 10)
ft <- flextable(dat)

ft <- color(ft, i = dat$Sepal.Length < 5, color = "orange")
ft
```

* j support also simple character vector containing the `col_key` names.

```{r}
dat <- head(iris, n = 10)
ft <- flextable(dat)
ft <- color(ft, j = "Sepal.Length", color = "orange", part = "all")
ft
```



