---
title: "as_flextable"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{as_flextable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>", 
  eval = TRUE
)
```

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(xtable)
library(officer)
library(flextable)
library(magrittr)
```

Function as_flextable is a generic function to cast objects into flextable.

So far, this method exist for objects `lm`, `glm`, `xtable`, `htest`, `as_grouped_data`.


## Groups as row titles


Package flextable does not support data transformation. A *grouped* dataset 
is then needed (and *tidy*). This has to be done prior to the creation of the object. 

Function `as_grouped_data` will modify data structure so that it will be 
easy to manage grouped data representation. Repeated consecutive values 
of group columns will be used to define the title of the groups and will 
be added as a *row title*.

Let's have an example with aggragated data from dataset `CO2`:

```{r}
library(data.table)
data_CO2 <- dcast(as.data.table(CO2), 
  Treatment + conc ~ Type, value.var = "uptake", fun.aggregate = mean)
head(data_CO2)
```

`as_grouped_data` will restructure the dataset:

```{r}
data_CO2 <- as_grouped_data(x = data_CO2, groups = c("Treatment"))
head(data_CO2)
```

The result is suitable for method `as_flextable`. A call to this function 
and few formatting operations are producing the following result:

```{r}
zz <- as_flextable( data_CO2 ) %>% 
  bold(j = 1, i = ~ !is.na(Treatment), bold = TRUE, part = "body" ) %>% 
  bold(part = "header", bold = TRUE ) %>% 
  width(width = 1.5)
zz
```

It's now easier to customize the rendering. Let's format column `conc` 
as an integer column:

```{r}
zz <- zz %>% 
  compose(i = ~ !is.na(conc), j = "conc", 
          value = as_paragraph(
            as_chunk(conc, formatter = function(x) sprintf("%.0f", x))
          )
  )
zz
```

Now let's add nice bars before displaying the figures:

```{r}
zz <- zz %>% 
  compose(i = ~ is.na(Treatment), j = "Quebec", 
          value = as_paragraph(
            minibar(Quebec), 
            " ", 
            as_chunk(Quebec, formatter = function(x) sprintf("%.01f", x))
            )
          ) %>% 
  compose(i = ~ is.na(Treatment), j = "Mississippi", 
          value = as_paragraph( minibar(Mississippi), 
                                " ",
                                as_chunk(Mississippi, 
                                         formatter = function(x) sprintf("%.01f", x) )
                                )
          ) %>% 
  align(j = 2:3, align = "left")
zz
```

And finally, add a footnote in the footer part:

```{r}
add_footer_lines(zz, "dataset CO2 has been used for this flextable") 
```

## xtable 

`xtable` objects can be transformed as flextable objects with function `as_flextable()`. 

```{r}
temp.ts <- ts(cumsum(1 + round(rnorm(100), 0)),
  start = c(1954, 7), frequency = 12)
ft <- as_flextable(x = xtable(temp.ts, digits = 0),
  NA.string = "-")
ft
```

### anova

```{r}
data(tli)
fm3 <- glm(disadvg ~ ethnicty*grade, data = tli, family = binomial)
ft <- as_flextable(xtable(anova(fm3)), hline.after = c(1))
ft
```

### adding horizontal lines

```{r}
bktbs <- xtable(matrix(1:10, ncol = 2))
hlines <- c(-1, 0, 1, nrow(bktbs))
ft <- as_flextable(bktbs, hline.after = hlines)
ft
```


### rotate columns 

```{r}
data(tli)
tli.table <- xtable(tli[1:10, ])
xtable::align(tli.table) <- "|r|r|clr|r|"
ft <- as_flextable(
  tli.table, 
  rotate.colnames = TRUE, 
  include.rownames = FALSE)

ft <- height(ft, i = 1, part = "header", height = 1)
ft
```

### tables

```{r}
Grade3 <- c("A","B","B","A","B","C","C","D","A","B",
  "C","C","C","D","B","B","D","C","C","D")
Grade6 <- c("A","A","A","B","B","B","B","B","C","C",
  "A","C","C","C","D","D","D","D","D","D")
Cohort <- table(Grade3, Grade6)
ft <- as_flextable(xtable(Cohort))
ft <- set_header_labels(ft, rowname = "Grade 3")
ft <- autofit(ft)
ft <- add_header(ft, A = "Grade 6") 
ft <- merge_at(ft, i = 1, j = seq_len( ncol(Cohort) ) + 1, 
  part = "header" ) 
ft <- bold(ft, j = 1, bold = TRUE, part = "body")
ft <- height_all(ft, part = "header", height = .4)
ft
```

### from scratch

```{r}
mat <- round(matrix(c(0.9, 0.89, 200, 0.045, 2.0), c(1, 5)), 4)
mat <- xtable(mat)
ft <- as_flextable(x = mat, NA.string = "-")
print(ft$col_keys)
superfp <- fp_text(vertical.align = "superscript", font.size = 8)

ft <- compose(ft, i = 1, j = "X1", part = "header", 
                  value = as_paragraph("R", as_chunk("2", props = superfp)) )
ft <- compose(ft, i = 1, j = "X2", part = "header", 
                  value = as_paragraph("\u03BC", as_chunk("x", props = superfp)))
ft <- compose(ft, i = 1, j = "rowname", part = "header", 
                  value = as_paragraph("y", as_chunk("t-1", props = superfp)))
ft <- set_header_labels(ft, X3 = "F-stat", X4 = "S.E.E", X5 = "DW", rowname = "")
ft <- autofit(ft)
ft
```

```{r echo=FALSE}
detach("package:xtable", unload = TRUE)
```

## Models and tests

`glm`, `lm` and `htest` objects can be easily converted to flextable:

### GLM example 

```{r}
clotting <- data.frame(
    u = c(5,10,15,20,30,40,60,80,100),
    lot1 = c(118,58,42,35,27,25,21,19,18),
    lot2 = c(69,35,26,21,18,16,13,12,12))
as_flextable(glm(lot1 ~ log(u), data = clotting, family = Gamma))
```

### LM example 

```{r}
ctl <- c(4.17,5.58,5.18,6.11,4.50,4.61,5.17,4.53,5.33,5.14)
trt <- c(4.81,4.17,4.41,3.59,5.87,3.83,6.03,4.89,4.32,4.69)
group <- gl(2, 10, 20, labels = c("Ctl","Trt"))
weight <- c(ctl, trt)
lm(weight ~ group) %>% as_flextable()
```

### Test example

```{r}
x <- rnorm(50)
y <- runif(30)
ks.test(x, y) %>% as_flextable()
```

