---
title: "flextable examples"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{flextable examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>", 
  eval = TRUE
)
```

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(officer)
library(flextable)
```




## Complex header

The following dataset is `typology`, a dataset containing data for table headers.

```{r echo=FALSE}
typology <- structure(list(col_keys = c("sep_1", "sep_2", "year", "premium",
"latest_eval", "cape_cod_u_l", "cape_cod_lr", "chain_ladder_u_l",
"chain_ladder_lr"), colC = c(" ", " ", "", "", "", "Cape Cod",
"Cape Cod", "Chain Ladder", "Chain Ladder"), colB = c(" ", " ",
"Year", "Premium", "Latest Eval", "Ultimate Loss", "Ultimate LR",
"Ultimate Loss", "Ultimate LR"), colA = c(" ", " ", "Year", "Premium",
"Latest Eval", "(000)", "(%)", "(000)", "(%)")), class = c("data.frame"), row.names = c(NA, -9L), .Names = c("col_keys",
"colC", "colB", "colA"))
typology
```

The following dataset is `x`, it will be displayed in the table body.


```{r echo=FALSE}
x <- structure(list(year = 2001:2004, premium = c(8.92042818214451,
12.6608266777834, 8.75775727892536, 9.85258034701945), latest_eval = c(4.49236487976272,
5.16555617550473, 6.22153711959364, 5.33407812665529), cape_cod_u_l = c(6998L,
7058L, 6923L, 6916L), cape_cod_lr = c(60L, 69L, 69L, 83L), chain_ladder_u_l = c(4.97029563327406,
5.98041693727634, 6.39257172913119, 4.40052968174841), chain_ladder_lr = c(69.3393557027241,
69.0607156360397, 71.4041411791844, 70.238484387929)), .Names = c("year",
"premium", "latest_eval", "cape_cod_u_l", "cape_cod_lr", "chain_ladder_u_l",
"chain_ladder_lr"), class = c("data.frame"), row.names = c(NA, -4L))
x
```



```{r}
double_format <- function(x){
  sprintf("%.3f", x)
}
percent_format <- function(x){
  sprintf("%.2f %%", x)
}
ft <- flextable(
  x, col_keys = c("year", "premium", "latest_eval",
                  "sep_1", "cape_cod_u_l", "cape_cod_lr",
                  "sep_2", "chain_ladder_u_l", "chain_ladder_lr") )
ft <- set_formatter(ft, premium = double_format, latest_eval = double_format,
                    chain_ladder_lr = percent_format )
ft <- set_header_df(ft, mapping = typology, key = "col_keys" )
ft <- merge_h(ft, part = "header")
ft <- merge_v(ft, part = "header", j = 1:3)
ft <- theme_zebra(ft, odd_header = "transparent", even_header = "transparent")

ft <- fontsize(ft, size = 12, part = "all")
ft <- color(ft, i = 1:2, color = "#007FA6", part = "header")
ft <- fontsize(ft, i = 3, size = 9, part = "header")
ft <- color(ft, i = 3, color = "gray", part = "header")

ft <- hline(ft, border = fp_border(width = .75, color = "#007FA6"), part = "body" )
ft <- hline(ft, border = fp_border(width = 2, color = "#007FA6"), part = "header" )

ft <- empty_blanks(ft)
ft <- autofit(ft)
ft
```

## Custom printing with knitr

To output data.frame as `flextable` in R Markdown documents, use the following 
code (and adapt if necessary):

```{r}
library(knitr)
dummy_fun = function(x, ...) {
  if( nrow(x) < 10 )
    knit_print(qflextable(x))
  else {
    ft <- qflextable(head(x, n = 15))
    ft <- add_footer_lines(ft, values = sprintf("(15 first lines /  %.0f lines)", nrow(x)))
    knit_print(theme_zebra(ft))
  }
}
# register the method
registerS3method("knit_print", "data.frame", dummy_fun)
```

```{r}
head(iris)
iris
```


## Looping in R Mardown documents

```{r}
library(htmltools)
ft <- flextable(head(iris))
tab_list <- list()
for(i in 1:3){
  tab_list[[i]] <- tagList(
    tags$h6(paste0("iteration ", i)),
    htmltools_value(ft)
  )
}
tagList(tab_list)
```

## Conditional formatting

Formatting functions accept arguments `i` and `j` to select rows 
and columns to format. These arguments support formulas, index, logical 
(and character for columns' names).

```{r}
ft <- flextable(head(mtcars))
ft <- color(ft, i = ~ drat > 3, j = ~ vs + am, color = "red")
ft <- bg(ft, i = ~ wt < 3, j = ~ mpg, bg = "#EFEF99")
ft <- bold(ft, i = 2:4, j = "cyl", bold = TRUE)
ft
```

## Compose paragraphs

```{r}
library(dplyr)
library(tidyr)

mydata <- data.frame(
  Espece = c("Acipenser Sturio (L. 1758) Esturgeon européen",
             "Alosa alosa  (L.1758) Alose vraie", "Alosa fallax (Lac. 1803) Alose feinte",
             "Anguilla anguilla  (L. 1758) Anguille", "Lampetra fluviatilis  (L. 1758) Lamproie de rivière",
             "Liza ramada (Risso 1826) Mulet porc") , 
  `X00` = c("",  "+", "+", "+", "",  "+"), 
  `X01` = c("",  "+", "+", "+", "+", "+")) %>% 
  separate(`Espece`, c("latin_name", "french_name"), sep = "\\(") %>% 
  mutate(french_name = paste0("(", french_name) )
mydata

flextable(mydata, col_keys = c("dummy", "X00", "X01")) %>%
  compose(j = "dummy", value = as_paragraph(as_i(latin_name), as_b(french_name))) %>% 
  color(color=c(
    rep("firebrick2",2),
    rep("black",3),
    rep("dodgerblue3",1)), 
    part="body") %>% 
  autofit()
```


## Correlation matrix

```{r}
library(flextable)
library(magrittr)

col_palette <- c("#D73027", "#F46D43", "#FDAE61", "#FEE08B", 
  "#D9EF8B", "#A6D96A", "#66BD63", "#1A9850")

cor_matrix <- cor(mtcars)

mycut <- cut(
  cor_matrix, 
  breaks = c(-1, -0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1), 
  include.lowest = TRUE, label = FALSE)

mycolors <- col_palette[mycut]
std_border <- fp_border(width = 1, color = "white")
data.frame(rowname = row.names(cor_matrix), stringsAsFactors = FALSE) %>%
  cbind(cor_matrix) %>%
  flextable() %>%
  border_outer(part="all", border = std_border ) %>% 
  border_inner_h(border = std_border, part="all") %>% 
  border_inner_v(border = std_border, part="all") %>% 
  compose(i = 1, j = 1, value = as_paragraph(""), part = "header") %>% 
  compose(j = colnames(cor_matrix), value = as_paragraph(""), part = "body") %>% 
  bg(j = colnames(cor_matrix), bg = mycolors) %>%
  align(align = "center", part = "all") %>%
  height_all(height = .5) %>% 
  width(width =  .5) 
  
```

## Count table

```{r}
library(data.table)
library(scales)
library(flextable)

Cancer <- fread("http://users.stat.ufl.edu/~aa/cat/data/Cancer.dat")
cancerCountWide <- dcast(Cancer, formula = time ~ histology + stage,
      fill = 0, value.var = "count", fun.aggregate = sum)
cancerCountWide

my_header <- data.frame(
  col_keys = c("time", "blank1", "1_1", "2_1", "3_1", 
               "blank2", "1_2", "2_2", "3_2", 
               "blank3", "1_3", "2_3","3_3", "label"),
  line2 = c("Follow-up", "", rep("I", 3), "", rep("II", 3), "", rep("III", 3), ""),
  line3 = c("Follow-up", rep(c("", "1", "2", "3"), 3), "")
)
my_header

colourer <- col_numeric(
  palette = c("transparent", "red"),
  domain = c(0, max(Cancer$count)+1))

flextable(cancerCountWide, col_keys = my_header$col_keys) %>% 
  set_header_df(mapping = my_header, key = "col_keys") %>% 
  theme_booktabs() %>% empty_blanks() %>% 
  flextable::compose(part = "header", j = "label", 
    value = as_paragraph(c("Histology", "Disease Stage"))) %>% 
  merge_v(part = "header", j = 1) %>% merge_h(part = "header", i = 1) %>% 
  align(align = "center", part = "all") %>% 
  align(align = "left", part = "header", j = "label") %>% 
  italic(part = "header", j = "label", italic = TRUE) %>% 
  bold(part = "header", bold = TRUE) %>% 
  bold(part = "body", bold = TRUE, j = 1) %>% 
  bg(
    bg = colourer, 
    j = c("1_1", "2_1", "3_1", "1_2", "2_2", 
          "3_2", "1_3", "2_3","3_3"), 
    part = "body") %>% 
  autofit() %>% 
  fix_border_issues()
```


## Using within shiny applications

Use function `htmltools_value()` to get the html value of the 
flextable (suitable for an `uiOutput`). 

```{r eval=FALSE}
library(shiny)
library(flextable)

ui <- fluidPage(
  
  titlePanel("mtcars"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("mpg", "mpg Limit", min = 11, max = 33, value = 20)
    ),
    mainPanel(
      uiOutput("mtcars_ft")
    )
  )
)

server <- function(input, output) {
  library(dplyr)
  output$mtcars_ft <- renderUI({
    req(input$mpg)
    mtcars %>%
      mutate(car = rownames(.)) %>%
      select(car, everything()) %>%
      filter(mpg <= input$mpg) %>%
      flextable() %>%
      theme_booktabs() %>% 
      htmltools_value()
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```

## Mixing links, footnotes and images 

I wrote this when I read the nice example written by Mara Averick with package `gt`, 

<blockquote class="twitter-tweet" data-lang="fr"><p lang="en" dir="ltr">😻 Aww, little tiny <a href="https://twitter.com/hashtag/NBA?src=hash&amp;ref_src=twsrc%5Etfw">#NBA</a> player photos…<br>{nbastatR} × {gt} 🙏 <a href="https://twitter.com/abresler?ref_src=twsrc%5Etfw">@abresler</a>  &amp; <a href="https://twitter.com/riannone?ref_src=twsrc%5Etfw">@riannone</a> <br>gt: Helper fun for adding images from the web: <a href="https://t.co/sbYgC4eZac">https://t.co/sbYgC4eZac</a><a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> code: <a href="https://t.co/174LemLojL">https://t.co/174LemLojL</a> <a href="https://t.co/Bfm8SOKkyC">pic.twitter.com/Bfm8SOKkyC</a></p>&mdash; Mara Averick (@dataandme) <a href="https://twitter.com/dataandme/status/1110531547168493571?ref_src=twsrc%5Etfw">26 mars 2019</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

I wanted to know what would be the flextable version:

```{r}
dat <- tibble::tribble(
  ~namePlayer, ~ratioBPM, ~ratioOBPM, ~ratioDBPM, ~urlPlayerStats, ~urlPlayerHeadshot,
"James Harden", 11.7, 10.5, 1.1, "https://stats.nba.com/player/201935", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/201935.png",
"Giannis Antetokounmpo", 10.8, 5.7, 5, "https://stats.nba.com/player/203507", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/203507.png",
"Nikola Jokic", 9.5, 5.7, 3.8, "https://stats.nba.com/player/203999", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/203999.png",
"Anthony Davis", 8.5, 4.7, 3.9, "https://stats.nba.com/player/203076", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/203076.png",
"LeBron James", 8.1, 6.2, 1.9, "https://stats.nba.com/player/2544", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/2544.png",
"Rudy Gobert", 7, 2, 5.1, "https://stats.nba.com/player/203497", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/203497.png",
"Karl-Anthony Towns", 6.8, 4.8, 2, "https://stats.nba.com/player/1626157", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/1626157.png",
"Russell Westbrook", 6.5, 2.5, 3.9, "https://stats.nba.com/player/201566", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/201566.png",
"Kyrie Irving", 6.4, 6, 0.4, "https://stats.nba.com/player/202681", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/202681.png",
"Nikola Vucevic", 6.4, 3, 3.4, "https://stats.nba.com/player/202696", "https://ak-static.cms.nba.com/wp-content/uploads/headshots/nba/latest/260x190/202696.png"
)
```

```{r}
ft <- flextable(dat, cwidth = c(1.7, .7, .7, .7, 1, .7)) %>%
  flextable::add_header_lines("Top 10 Box Plus/Minus") %>%
  set_header_labels(namePlayer = "", ratioBPM = "BPM", ratioOBPM = "OBPM",
    ratioDBPM = "DBPM", urlPlayerStats = "Link", urlPlayerHeadshot = "") %>%
  compose( j = "urlPlayerStats",
           value = as_paragraph( hyperlink_text(x = "player stats", url = urlPlayerStats ) ) ) %>% 
  compose( j = "urlPlayerHeadshot",
           value = as_paragraph(
             as_image(src = urlPlayerHeadshot, width = .52, height = .38)
           ), part = "body") %>% 
  footnote( i = 1, j = 1, ref_symbols = "*",
            value = as_paragraph("Players with 500+ minutes."), part = "header") %>% 
  footnote( i = 2, j = 2:4, ref_symbols = c("†", "‡", "§"),
            value = as_paragraph(
              c("Box Plus/Minus: a box score estimate of the points per 100 possessions that a player contributed above a league-average player, translated to an average team.", 
                "Offensive Box Plus/Minus.", "Defensive Box Plus/Minus.")
            ), part = "header") %>% 
  theme_zebra(odd_header="transparent") %>% 
  valign(valign = "bottom", part = "all") %>% 
  fontsize(i = 1, part = "header", size = 20) %>% 
  bold(i = 1, part = "header", bold = TRUE) %>% 
  bold(part = "footer", bold = FALSE) %>% 
  italic(part = "footer", italic = TRUE) %>% 
  color(j = 5, color = "#337ab7") %>% 
  vline(j = 1, part = "body", border = officer::fp_border()) %>% 
  border_inner_h(part = "header", border = officer::fp_border()) %>% 
  hline_bottom(part = "all", border = officer::fp_border()) %>% 
  colformat_num(col_keys = c("ratioBPM", "ratioOBPM", "ratioDBPM"), digits = 1 )

ft
```

