---
title: "Render as image"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Render as image}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>", 
  eval = TRUE
)
```

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(officer)
library(flextable)
```

Beeing able to render flextable as images is requiring package `webshot` 
or `webshot2` to be available.

Run the following command to install `webshot2`. This package should be used 
instead of `webshot` as the latest can have some issues with html rendering.

```{r eval=FALSE}
devtools::install_github("rstudio/webshot2")
```

> Read more about it [here](https://github.com/rstudio/webshot2).

Run the following command to install `webshot`. You may have to install `phantomjs` 
first.

```{r eval=FALSE}
install.packages("webshot")
```

## What will be the size?

You can print a table as an image. In some case, you may want to know the exact width and height 
of the table to be sure to set up the correct aspect ratio. The function `flextable_dim()` will 
provide these informations without the need to produce the image.

```{r}
ft <- flextable(head(iris))
dims <- flextable_dim(ft)
dims
```

Results are in inches. You can reuse them as values for knitr chunk options `fig.asp`, `fig.width` and `fig.height`.


## Images

To get an image file from a `flextable` object, use function `save_as_image`.

```{r}
img_file <- tempfile(fileext = ".png")
save_as_image(ft, path = img_file, webshot = "webshot2")
```


## As an R graphic

To get a plot from a `flextable` object, use function `plot`.

```{r fig.asp=dims$aspect_ratio, fig.width=dims$width*3,out.width=dims$width*72}
plot(ft)
```

> To add a flextable on a ggplot2 object, one can use `flextable::as_raster` 
and `grid.raster` or `rasterGrob`. See [Examples](./examples.html#send-your-flextable-to-ggplot) for an example.

## With ggplot 

You can also add the flextable (with `grid::rasterGrob()` and `ggplot2::annotation_custom()`) 
in a ggplot object. You need first to use `as_raster()` function that will return a raster 
version of the flextable.

```{r out.width="70%", fig.width=8, fig.height=5}
library(ggplot2)
library(grid)
library(cowplot)
library(dplyr)

gg1 <- ggplot(iris, aes(Sepal.Length, Petal.Length, color = Species) ) + geom_point()

ft_raster <- iris %>% group_by(Species) %>% 
  summarise_all(median) %>% 
  flextable() %>% autofit() %>% add_header_lines("means per Species") %>% 
  as_raster(webshot = "webshot2")

gg2 <- ggplot() + 
  theme_void() + 
  annotation_custom(rasterGrob(ft_raster), xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)

cowplot::plot_grid(gg1, gg2, nrow = 2, ncol = 1, rel_heights = c(4, 1) )
```
